### 2024-11-23 paper/paper.qmd
do data visualization

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - First author
  - Another author
thanks: "Code and data are available at: [https://github.com/RohanAlexander/starter_folder](https://github.com/RohanAlexander/starter_folder)."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(ggplot2)
library(dplyr)
library(here)
```

```{r}
data <- read_csv(here("data", "01-analysis_data", "analysis_data.csv"))
```

# Introduction

Overview paragraph

Estimand paragraph

Results paragraph

Why it matters paragraph

Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....

# Data {#sec-data}

## Overview

We use the statistical programming language R [@citeR].... Our data [@shelter].... Following @tellingstories, we consider...

Overview text

## Measurement

Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.

## Outcome variables

Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.

```{r}

```

Talk way more about it.

```{r}
#| echo: false
#| warning: false
#| label: fig-demographics-distribution
#| fig-cap: "Demographics Distribution and Average Income by Group"
#| fig-subcap: [
#| "Gender Proportion", "Average Income by Gender", 
#| "Race Proportion", "Average Income by Race", 
#| "Region Proportion", "Average Income by Region", 
#| "Education Proportion", "Average Income by Education Level"
#| ]
#| layout-ncol: 2

# Gender proportion and income
gender_count <- data %>%
  group_by(gender) %>%
  summarise(count = n()) %>%
  mutate(percentage = round(count / sum(count) * 100, 1))

pie_chart_gender <- ggplot(gender_count, aes(x = "", y = percentage, fill = gender)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y") +
  geom_text(aes(label = paste0(percentage, "%")),
            position = position_stack(vjust = 0.5), size = 3) +
  theme_void() +
  theme(legend.position = "right") +
  labs(fill = "Gender")

avg_income_gender <- data %>%
  group_by(gender) %>%
  summarise(Average_Income = mean(INCWAGE, na.rm = TRUE))

bar_chart_gender <- ggplot(avg_income_gender, aes(x = gender, y = Average_Income, fill = gender)) +
  geom_bar(stat = "identity", width = 0.6) +
  geom_text(aes(label = round(Average_Income, 0)), 
            vjust = -0.5, size = 3) +
  labs(x = "Gender", y = "Average Income") +
  theme_minimal() +
  scale_y_continuous(expand = expansion(mult = c(0, 0.2))) +  # Expand upper limit
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "none"
  )

# Race proportion and income
race_count <- data %>%
  group_by(race_group) %>%
  summarise(count = n()) %>%
  mutate(percentage = round(count / sum(count) * 100, 1))

pie_chart_race <- ggplot(race_count, aes(x = "", y = percentage, fill = race_group)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y") +
  geom_text(aes(label = paste0(percentage, "%")),
            position = position_stack(vjust = 0.5), size = 3) +
  theme_void() +
  theme(legend.position = "right") +
  labs(fill = "Race Group")

avg_income_race <- data %>%
  group_by(race_group) %>%
  summarise(Average_Income = mean(INCWAGE, na.rm = TRUE))

bar_chart_race <- ggplot(avg_income_race, aes(x = race_group, y = Average_Income, fill = race_group)) +
  geom_bar(stat = "identity", width = 0.6) +
  geom_text(aes(label = round(Average_Income, 0)), 
            vjust = -0.5, size = 3) +
  labs(x = "Race Group", y = "Average Income") +
  theme_minimal() +
  scale_y_continuous(expand = expansion(mult = c(0, 0.2))) +  # Expand upper limit
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "none"
  )

# Region proportion and income
region_count <- data %>%
  group_by(region) %>%
  summarise(count = n()) %>%
  mutate(percentage = round(count / sum(count) * 100, 1))

pie_chart_region <- ggplot(region_count, aes(x = "", y = percentage, fill = region)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y") +
  geom_text(aes(label = paste0(percentage, "%")),
            position = position_stack(vjust = 0.5), size = 3) +
  theme_void() +
  theme(legend.position = "right") +
  labs(fill = "Region")

avg_income_region <- data %>%
  group_by(region) %>%
  summarise(Average_Income = mean(INCWAGE, na.rm = TRUE))

bar_chart_region <- ggplot(avg_income_region, aes(x = region, y = Average_Income, fill = region)) +
  geom_bar(stat = "identity", width = 0.6) +
  geom_text(aes(label = round(Average_Income, 0)), 
            vjust = -0.5, size = 3) +
  labs(x = "Region", y = "Average Income") +
  theme_minimal() +
  scale_y_continuous(expand = expansion(mult = c(0, 0.2))) +  # Expand upper limit
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "none"
  )
# Education proportion and income
education_count <- data %>%
  group_by(education_level) %>%
  summarise(count = n()) %>%
  mutate(percentage = round(count / sum(count) * 100, 1))

# 设置学历顺序
education_levels <- c("Below_High_School", "High_School", "Some_College", "Bachelor", "Above_Bachelor")
education_count <- education_count %>%
  mutate(education_level = factor(education_level, levels = education_levels))

avg_income_education <- data %>%
  group_by(education_level) %>%
  summarise(Average_Income = mean(INCWAGE, na.rm = TRUE)) %>%
  mutate(education_level = factor(education_level, levels = education_levels))

# Pie chart: education level proportion
pie_chart_education <- ggplot(education_count, aes(x = "", y = percentage, fill = education_level)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y") +
  geom_text(aes(label = paste0(percentage, "%")),
            position = position_stack(vjust = 0.5), size = 3) +
  theme_void() +
  theme(legend.position = "right") +
  labs(fill = "Education Level")

# Bar chart: average income by education level, ordered
bar_chart_education <- ggplot(avg_income_education, aes(x = education_level, y = Average_Income, fill = education_level)) +
  geom_bar(stat = "identity", width = 0.6) +
  geom_text(aes(label = round(Average_Income, 0)), 
            vjust = -0.5, size = 3) +
  labs(x = "Education Level", y = "Average Income") +
  theme_minimal() +
  scale_y_continuous(expand = expansion(mult = c(0, 0.2)), labels = scales::comma) +  # Use comma format for y-axis
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "none"
  )
# Output all eight charts in four rows with two charts per row
pie_chart_gender
bar_chart_gender
pie_chart_race
bar_chart_race
pie_chart_region
bar_chart_region
pie_chart_education
bar_chart_education

```

## Predictor variables

Add graphs, tables and text.

Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

Our results are summarized in @tbl-modelresults.

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

Please don't use these as sub-heading labels - change them to be what your point actually is.

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

\newpage

# References

### 2024-11-23 scripts/00-simulate_data.R

@@ -29,27 +29,27 @@ region <- sample(c("Northeast", "South", "Midwest", "West"), n_samples, replace 
 
 # Simulate detailed education levels
 education_level <- sample(
-  c("No_Education", "Primary_Education", "Middle_School", "High_School", 
-    "Some_College", "Bachelors_or_Higher"),
+  c("Below_High_School", "High_School", "Some_College", "Bachelor", "Above_Bachelor"),
   n_samples, 
   replace = TRUE, 
-  prob = c(0.05, 0.1, 0.15, 0.4, 0.2, 0.1)
+  prob = c(0.25, 0.4, 0.2, 0.1, 0.05)
 )
 
 # Simulate working hours based on education level
 uhrswork <- rnorm(n_samples, mean = 40, sd = 5)
-uhrswork[education_level == "No_Education"] <- rnorm(sum(education_level == "No_Education"), mean = 25, sd = 5)
-uhrswork[education_level == "Primary_Education"] <- rnorm(sum(education_level == "Primary_Education"), mean = 30, sd = 5)
-uhrswork[education_level == "Middle_School"] <- rnorm(sum(education_level == "Middle_School"), mean = 35, sd = 5)
+uhrswork[education_level == "Below_High_School"] <- rnorm(sum(education_level == "Below_High_School"), mean = 30, sd = 5)
+uhrswork[education_level == "High_School"] <- rnorm(sum(education_level == "High_School"), mean = 35, sd = 5)
+uhrswork[education_level == "Some_College"] <- rnorm(sum(education_level == "Some_College"), mean = 38, sd = 5)
+uhrswork[education_level == "Bachelor"] <- rnorm(sum(education_level == "Bachelor"), mean = 42, sd = 5)
+uhrswork[education_level == "Above_Bachelor"] <- rnorm(sum(education_level == "Above_Bachelor"), mean = 45, sd = 5)
 
 # Simulate income based on education, region, and gender with interaction effects
 education_to_income <- c(
-  "No_Education" = 15000,
-  "Primary_Education" = 20000,
-  "Middle_School" = 25000,
+  "Below_High_School" = 20000,
   "High_School" = 35000,
   "Some_College" = 45000,
-  "Bachelors_or_Higher" = 65000
+  "Bachelor" = 60000,
+  "Above_Bachelor" = 80000
 )
 base_income <- as.numeric(education_to_income[education_level])
 region_factor <- ifelse(region == "Northeast", 1.1, ifelse(region == "South", 0.9, 1.0))
[…]

### 2024-11-23 scripts/01-test_simulated_data.R

@@ -74,8 +74,7 @@ test_that("Value ranges and logical consistency", {
   expect_true(all(simulated_data$region %in% valid_regions), "Invalid values found in region.")
   
   # Test 16: education_level contains valid categories
-  valid_education <- c("No_Education", "Primary_Education", "Middle_School", 
-                       "High_School", "Some_College", "Bachelors_or_Higher")
+  valid_education <- c("Below_High_School", "High_School", "Some_College", "Bachelor", "Above_Bachelor")
   expect_true(all(simulated_data$education_level %in% valid_education), "Invalid values found in education_level.")
   
   # Test 17: gender contains valid categories
@@ -83,11 +82,10 @@ test_that("Value ranges and logical consistency", {
   expect_true(all(simulated_data$gender %in% valid_genders), "Invalid values found in gender.")
   
   # Test 18: race_group contains valid categories
-  valid_race_groups <- c("White", "Black", "Asian", "Native", "Mixed_Other")
+  valid_race_groups <- c("White", "Black", "Asian", "Other")
   expect_true(all(simulated_data$race_group %in% valid_race_groups), "Invalid values found in race_group.")
 })
 
-  
 
 
 

### 2024-11-23 scripts/03-clean_data.R

@@ -35,12 +35,11 @@ cleaned_data <- raw_data %>%
   filter(!is.na(region)) %>%
   filter(!EDUC %in% c(0, 99)) %>%  # Remove invalid education values
   mutate(education_level = case_when(
-    EDUC == 1 ~ "No_Education",
-    EDUC == 2 ~ "Primary_Education",
-    EDUC == 3 ~ "Middle_School",
-    EDUC %in% c(4, 5, 6) ~ "High_School",
-    EDUC %in% c(7, 8, 9) ~ "Some_College",
-    EDUC %in% c(10, 11) ~ "Bachelors_or_Higher"
+    EDUC %in% c(0, 1, 2, 3, 4, 5) ~ "Below_High_School",  
+    EDUC == 6 ~ "High_School",  
+    EDUC %in% c(7, 8) ~ "Some_College",  
+    EDUC == 10 ~ "Bachelor",  
+    EDUC == 11 ~ "Above_Bachelor"  
   )) %>%
   mutate(age = ifelse(AGE == 999, NA, AGE)) %>%  # Replace 999 with NA
   filter(!is.na(age) & age >= 18 & age <= 65) %>%  # Keep working age (18-65)
@@ -54,9 +53,8 @@ cleaned_data <- raw_data %>%
   mutate(race_group = case_when(
     RACE == 1 ~ "White",
     RACE == 2 ~ "Black",
-    RACE == 3 ~ "Native",
-    RACE %in% c(4, 5, 6) ~ "Asian",
-    RACE %in% c(7, 8, 9) ~ "Mixed_Other"
+    RACE %in% c(3, 7, 8, 9) ~ "Other",  
+    RACE %in% c(4, 5, 6) ~ "Asian"
   )) %>%
   filter(EMPSTAT == 1) %>%  # Keep only employed individuals
   select(-STATEFIP, -RACE, -EDUC, -EMPSTAT, -SEX, -AGE)

### 2024-11-23 scripts/04-test_analysis_data.R

@@ -75,8 +75,7 @@ test_that("Value ranges and logical consistency", {
   expect_true(all(analysis_data$region %in% valid_regions), "Invalid values found in region.")
   
   # Test 16: education_level contains valid categories
-  valid_education <- c("No_Education", "Primary_Education", "Middle_School", 
-                       "High_School", "Some_College", "Bachelors_or_Higher")
+  valid_education <- c("Below_High_School", "High_School", "Some_College", "Bachelor", "Above_Bachelor")
   expect_true(all(analysis_data$education_level %in% valid_education), "Invalid values found in education_level.")
   
   # Test 17: gender contains valid categories
@@ -84,7 +83,7 @@ test_that("Value ranges and logical consistency", {
   expect_true(all(analysis_data$gender %in% valid_genders), "Invalid values found in gender.")
   
   # Test 18: race_group contains valid categories
-  valid_race_groups <- c("White", "Black", "Asian", "Native", "Mixed_Other")
+  valid_race_groups <- c("White", "Black", "Asian", "Other")
   expect_true(all(analysis_data$race_group %in% valid_race_groups), "Invalid values found in race_group.")
   
   # Test 19: Logical consistency: Individuals with no UHRSWORK should have zero income

### 2024-11-23 paper/paper.qmd
add some graph

@@
 library(here)
+library(reshape2)
 ```
@@
 
+```{r}
+#| echo: false
+#| warning: false
+#| label: fig-Education-Race
+#| fig-cap: "Proportion of Education Levels by Race Group"
+
+# Ensure the order of education levels, arranged from low to high
+education_order <- c("Below_High_School", "High_School", "Some_College", "Bachelor", "Above_Bachelor")
+race_order <- c(unique(data$race_group[data$race_group != "Other"]), "Other")
+
+# Set education levels and race groups as factors with specified order
+data$education_level <- factor(data$education_level, levels = education_order)
+data$race_group <- factor(data$race_group, levels = race_order)
+
+# Create a contingency table and fill missing values with 0
+education_race_crosstab <- prop.table(table(data$education_level, data$race_group), margin = 2)
+education_race_crosstab[is.na(education_race_crosstab)] <- 0  # Fill NaN with 0
+
+# Convert the data from wide to long format
+heatmap_data <- melt(education_race_crosstab)
+
+# Plot the heatmap with proportion labels and dynamic text color
+ggplot(heatmap_data, aes(Var2, Var1, fill = value)) +
+  geom_tile(color = "white") +  # Heatmap tiles
+  geom_text(
+    aes(
+      label = sprintf("%.2f", value),
+      color = ifelse(value > 0.25, "white", "black")  
+    ),
+    size = 4
+  ) +
+  scale_fill_gradient(low = "white", high = "blue", name = "Proportion") +
+  scale_color_manual(values = c("black" = "black", "white" = "white"), guide = "none") +  # Define color mapping
+  labs(
+    x = "Race Group",
+    y = "Education Level"
+  ) +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+
+```
+
 ## Predictor variables

### 2024-11-24 paper/paper.qmd
add data result

@@
 library(reshape2)
+library(scales)
+library(arrow)
+library(knitr)
 ```
@@
 ```{r}
-data <- read_csv(here("data", "01-analysis_data", "analysis_data.csv"))
+#| echo: false
+#| warning: false
+#| message: false
+
+data <- read_parquet(here("data", "01-analysis_data", "analysis_data.parquet"))
+
 ```
@@
 
-## Overview
+## Data Overview
+Our data [@shelter]....
 
-We use the statistical programming language R [@citeR].... Our data [@shelter].... Following @tellingstories, we consider...
+We use the statistical programming language R [@citeR]....Our data [@shelter]....
 
-Overview text
+The IPUMS USA database [@ipums2024] is one of the largest collections of microdata from population censuses globally.
+Supported by organizations such as the National Institutes of Health and the University of Minnesota, it includes data from the American Community Survey (ACS) and other census programs.
+This resource provides detailed individual-level data, allowing researchers to address specific social, economic, and demographic questions.
+A key feature of IPUMS USA is its capacity to create customized datasets by selecting variables based on research needs, minimizing the inclusion of irrelevant data.
+
+For this study, data from the 2023 ACS sample in IPUMS USA were used.
+The selected variables include STATEFIP (state code), SEX (gender), AGE (age), RACE (race), EDUC (educational attainment), EMPSTAT (employment status), UHRSWORK (usual hours worked per week), and INCWAGE (wage and salary income).
+These variables cover geographic location, demographic characteristics, education, and employment conditions, providing a broad basis for examining factors influencing wage income.
+
+Other databases, such as IPUMS CPS [@ipums_cps2024] and the National Longitudinal Surveys [@nlsy1979], were considered but deemed less suitable for this analysis.
+IPUMS CPS provides detailed labor market data but has a smaller sample size and limited geographic detail.
+The National Longitudinal Surveys track specific population groups over time but lack broad representation, making them less ideal for cross-sectional studies.
+In contrast, the 2023 ACS data in IPUMS USA offers larger population coverage and a diverse range of variables, making it a better choice for analyzing wage determinants.
 
@@
 
-Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.
+This study utilizes sample data from the 2023 ACS.
+The dataset is based on a 1% random national sample, ensuring broad representativeness.
+It includes individuals living in private households as well as those in group quarters, such as student dormitories and care facilities.
+Weights were applied to the data to account for potential biases introduced by the sampling design and nonresponse.
 
-## Outcome variables
+Data collection for the ACS involved structured questionnaires covering a wide range of topics, such as demographics, education, employment, and income.
+For example, UHRSWORK (usual hours worked per week) captures labor input based on respondents’ reports of their average weekly work hours.
+The smallest geographic unit in the dataset is the Public Use Microdata Area (PUMA), which contains at least 100,000 residents and falls within state boundaries.
+This structure protects respondent privacy while supporting regional analysis.
 
-Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.
+While the ACS methodology ensures data reliability and extensive coverage, variables like income and work hours rely on self-reported responses, which may introduce minor inaccuracies.
+Nonetheless, the use of sampling weights and robust design make the dataset a reliable foundation for studying wage income and its associated factors.
+A detailed explanation of the survey methodology and sampling design is provided in the appendix.
+
+## Dataset Description
 
 ```{r}
+#| echo: false
+#| warning: false
+
+# 确保每个分类变量的所有类别都被显示
+complete_data <- data %>%
+  group_by(education_level) %>%  # 替换为需要展示的分类变量
+  summarise_all(~ first(.))  # 保证每个类别至少展示一行
+
+# 使用 kable 生成表格
+library(knitr)
+kable(complete_data, caption = "Table 1: Categories of SEX Fully Displayed", align = 'c')
 
@@
 
-Talk way more about it.
+The cleaned dataset includes the following variables.
+UHRSWORK and INCWAGE are retained from the raw data, while the other variables are newly constructed.
+During the data cleaning process, only individuals aged 18-65 and those with an employment status of "Employed" were included.
+This ensures the dataset focuses on studying the wage determinants of the employed labor force.
+Below is a description of each variable and how it was constructed:
+
+UHRSWORK: Reports the number of hours per week the respondent usually worked if they were employed during the reference period.
+
+INCWAGE: Represents each respondent's total pre-tax wage and salary income earned as an employee during the previous year.
+
+region: Constructed from the STATEFIP variable to represent geographic regions.
+Valid state codes (1–56) were grouped into four traditional U.S.
+regions: Northeast, Midwest, South, and West.
+Each state was assigned to its corresponding region for further regional analysis.
+
+education_level: Derived from the EDUC variable, which indicates respondents' educational attainment as measured by the highest year of school or degree completed.
+It is grouped into five categories: Below High School (includes all levels below high school), High School (completed grade 12), Some College (completed 1-2 years of college), Bachelor’s Degree (completed 4 years of college), and Above Bachelor (more than 5 years of higher education).age: Filtered from the AGE variable to include only respondents aged 18 to 65.
+This variable records the respondent’s age in years as of their last birthday.
+
+gender: Based on the SEX variable, with original codes recoded as “Male” and “Female.”
+
+race_group: Constructed from the RACE variable and grouped into four categories: White, Black, Asian, and Other.
+White and Black remain consistent with the original classifications.
+Asian includes Chinese, Japanese, and Other Asian categories, while Other includes all other races and mixed-race groups.
+
+## Data Results
 
@@
 
-# 设置学历顺序
 education_levels <- c("Below_High_School", "High_School", "Some_College", "Bachelor", "Above_Bachelor")
@@
 
+@fig-demographics-distribution illustrates the distribution of demographic characteristics and their relationship with average income across gender, race, region, and education level.
+Pie charts display the proportional distribution of each group, while bar charts show corresponding average income levels, providing a straightforward depiction of these relationships.
+
+Chart (a) shows the gender distribution: females account for 48.9% of the population, and males 51.1%, indicating a nearly equal split.
+Chart (b) shows that males earn an average income of $82,840, compared to $57,146 for females.
+
+Chart (c) presents the racial composition, with White individuals forming the largest group at 67%, followed by Black individuals at 18.2%, Asians at 6.8%, and other racial groups at 8.1%.
+Chart (d) shows that Asians have the highest average income at $96,574, followed by White individuals at $73,778.
+Black individuals and other racial groups earn $51,221 and $56,604, respectively.
+
+Chart (e) shows the regional distribution.
+The West represents the largest share at 29%, followed by the Northeast (26.6%), the South (24.1%), and the Midwest (20.3%).
+Chart (f) indicates that the Midwest has the lowest average income at $62,046, while the Northeast has the highest at $78,138.
+The South and West have average incomes of $66,954 and $75,293, respectively.
+
+Chart (g) shows education levels: 16.5% of the population has less than a high school education, 30.4% has completed high school, 23.5% has some college education, 26.2% holds a bachelor’s degree, and 9.5% has education beyond a bachelor’s degree.
+Chart (h) highlights the connection between education and income.
+Those with less than a high school education earn $36,149 on average, while individuals with a bachelor’s degree or higher earn $87,465 and $124,144, respectively.
+Income increases steadily with higher levels of education.
+
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+#| label: fig-combined-visualizations
+#| fig-cap: "Income Trends by Age and Work Hours"
+#| fig-subcap: [
+#| "Work Hours vs Income", 
+#| "Average Income by Age", 
+#| "Age vs Income", 
+#| "Average Income by Work Hours"
+#| ]
+#| layout-ncol: 2
+
+# Calculate average income by work hours
+workhours_avg_income <- data %>%
+  group_by(UHRSWORK) %>%
+  summarise(avg_income = mean(INCWAGE, na.rm = TRUE))
+
+# Calculate average income by age
+age_avg_income <- data %>%
+  group_by(age) %>%
+  summarise(avg_income = mean(INCWAGE, na.rm = TRUE))
+
+# Work Hours vs Income with quadratic fit
+plot1 <- ggplot(data, aes(x = UHRSWORK, y = INCWAGE)) +
+  geom_point(color = "blue", alpha = 0.6) +
+  geom_smooth(method = "lm", formula = y ~ poly(x, 2), color = "red") +
+  labs(x = "Work Hours", y = "Income") +
+  theme_minimal() +
+  theme(plot.title = element_blank()) +
+  scale_y_continuous(labels = comma)
+
+# Line chart of average income by work hours 
+plot2 <- ggplot(workhours_avg_income, aes(x = UHRSWORK, y = avg_income)) +
+  geom_line(color = "blue", size = 1) +
+  geom_point(color = "blue", size = 2) +
+  labs(x = "Work Hours", y = "Average Income") +
+  theme_minimal() +
+  theme(plot.title = element_blank())
+
+# Age vs Income with quadratic fit
+plot3 <- ggplot(data, aes(x = age, y = INCWAGE)) +
+  geom_point(color = "orange", alpha = 0.6) +
+  geom_smooth(method = "lm", formula = y ~ poly(x, 2), color = "red") +
+  labs(x = "Age", y = "Income") +
+  theme_minimal() +
+  theme(plot.title = element_blank()) +
+  scale_y_continuous(labels = comma)
+
+# Line chart of average income by age (no change needed)
+plot4 <- ggplot(age_avg_income, aes(x = age, y = avg_income)) +
+  geom_line(color = "orange", size = 1) +
+  geom_point(color = "orange", size = 2) +
+  labs(x = "Age", y = "Average Income") +
+  theme_minimal() +
+  theme(plot.title = element_blank())
+
+# Display the plots
+plot1
+plot2
+plot3
+plot4
+
+```
+
+@fig-combined-visualizations examines the relationship between income, work hours, and age, incorporating both individual-level data and aggregated trends.
+
+Chart (a) is a scatter plot of work hours versus income, with each point representing an individual.
+The red quadratic regression line indicates a positive correlation between income and work hours, where longer hours generally lead to higher income.
+However, the regression line primarily reflects a linear trend, suggesting a stable relationship between income and work hours without significant signs of deceleration.
+Additionally, income variability increases with longer work hours, particularly beyond 50 hours per week, where fluctuations become more pronounced.
+Chart (b) illustrates the relationship between work hours and average income.
+While income generally increases with work hours, especially between 20 and 50 hours per week, it becomes unstable beyond 50 hours.
+
+Chart (c) is a scatter plot of age versus income, with the red quadratic regression line showing that income increases with age before eventually declining.
+Chart (d) depicts the trend of average income by age.
+The line chart shows that average income steadily rises with age, peaking in the late 40s and early 50s, and gradually declines as individuals approach retirement age.
+
 ```{r}
@@
 
-## Predictor variables
+The chart illustrates the proportion of different education levels across racial groups.
+The horizontal axis represents racial groups, including "Asian," "Black," "White," and "Other," while the vertical axis represents education levels, ranging from "Below High School" to "Above Bachelor." The intensity of blue indicates the proportion, with darker shades representing higher values.
 
-Add graphs, tables and text.
+Among Asians, the proportion of highly educated individuals is relatively high, with 38% holding a bachelor's degree and 30% having education beyond a bachelor's degree, while the proportion of those with "Below High School" education is 0%.
+In the Black group, "High School" accounts for the largest proportion at 32%, followed by "Some College" at 27%, while those with "Above Bachelor" degrees make up only 16%.
+In the White group, "High School" accounts for 29%, with 27% holding a bachelor's degree and 24% having "Some College" education, while 17% have "Above Bachelor" degrees.
+Among the "Other" group, 40% have a "High School" education, 25% have "Some College," 17% hold a bachelor's degree, and only 9% have "Above Bachelor" degrees.
 
-Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.
+Overall, the distribution of education levels varies across racial groups.
+Asians have a higher proportion of highly educated individuals compared to other groups, while "High School" education is more prevalent among Black and "Other" groups.
+This reflects the complex relationship between race and educational background.
 

### 2024-11-24 scripts/03-clean_data.R

@@ -5,7 +5,7 @@
 # Contact: lijianing.li@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: 
-# - The `tidyverse`, `dplyr` and `here` packages must be installed and loaded.
+# - The `tidyverse`, `dplyr`，`arrow` and `here` packages must be installed and loaded.
 # - The steps outlined in 02-download_data.R this document must be followed.
 # Any other information needed? Make sure you are in the `Determinants-of-Wages` rproj
 
@@ -13,6 +13,7 @@
 library(tidyverse)
 library(dplyr)
 library(here)
+library(arrow)
 
 #### Clean data ####
 # Read data
@@ -65,6 +66,6 @@ analysis_data <- cleaned_data %>%
   slice_sample(n = 5000)
 
 # Save the cleaned data
-output_path <- here("data", "01-analysis_data", "analysis_data.csv")
-write.csv(analysis_data, output_path, row.names = FALSE)
+output_path <- here("data", "01-analysis_data", "analysis_data.parquet")
+write_parquet(analysis_data, output_path)
 

### 2024-11-24 scripts/04-test_analysis_data.R

@@ -5,7 +5,7 @@
 # Contact: lijianing.li@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: 
-# - The `tidyverse`, `testthat`, `readr` and `here` packages must be installed and loaded.
+# - The `tidyverse`, `testthat`, `readr`，`arrow` and `here` packages must be installed and loaded.
 # - 03-clean_data.R must have been run
 # Any other information needed? Make sure you are in the `Determinants-of-Wages` rproj
 
@@ -14,8 +14,9 @@ library(tidyverse)
 library(testthat)
 library(readr)
 library(here)
+library(arrow)
 
-analysis_data <- read_csv(here("data", "01-analysis_data", "analysis_data.csv"))
+analysis_data <- read_parquet(here("data", "01-analysis_data", "analysis_data.parquet"))
 
 #### Test data ####
 # Tests for Dataset Structure and Loading
@@ -38,10 +39,10 @@ test_that("Dataset structure and loading", {
 # Tests for Column Data Types
 test_that("Column data types", {
   # Test 5: UHRSWORK is numeric
-  expect_type(analysis_data$UHRSWORK, "double")
+  expect_type(analysis_data$UHRSWORK, "integer")
   
   # Test 6: INCWAGE is numeric
-  expect_type(analysis_data$INCWAGE, "double")
+  expect_type(analysis_data$INCWAGE, "integer")
   
   # Test 7: region is a character vector
   expect_type(analysis_data$region, "character")
@@ -50,7 +51,7 @@ test_that("Column data types", {
   expect_type(analysis_data$education_level, "character")
   
   # Test 9: age is numeric
-  expect_type(analysis_data$age, "double")
+  expect_type(analysis_data$age, "integer")
   
   # Test 10: gender is a character vector
   expect_type(analysis_data$gender, "character")

### 2024-11-25 paper/paper.qmd
add model

@@
 library(knitr)
+library(kableExtra)
 ```
@@
 #| echo: false
+#| eval: true
+#| label: tbl-datasummary
+#| tbl-cap: "Cleaned Data Overview"
 #| warning: false
+#| results: asis
 
-# 确保每个分类变量的所有类别都被显示
+# Ensure all categories of each categorical variable are displayed
 complete_data <- data %>%
-  group_by(education_level) %>%  # 替换为需要展示的分类变量
-  summarise_all(~ first(.))  # 保证每个类别至少展示一行
+  group_by(education_level) %>%      
+  summarise_all(~ first(.)) %>%       # Ensure at least one row is displayed for each category
+  select(INCWAGE, UHRSWORK, everything())
 
-# 使用 kable 生成表格
-library(knitr)
-kable(complete_data, caption = "Table 1: Categories of SEX Fully Displayed", align = 'c')
-
+kable(
+  complete_data, 
+  align = 'c',
+  format = "latex"
+) %>%
+  kable_styling(
+    latex_options = "scale_down",  
+    full_width = FALSE
+  ) %>%
+  column_spec(1, width = "3cm")          
 ```
 
-The cleaned dataset includes the following variables.
+The sample of the cleaned dataset is shown in @tbl-datasummary, including the following variables.
 UHRSWORK and INCWAGE are retained from the raw data, while the other variables are newly constructed.
@@
 
-UHRSWORK: Reports the number of hours per week the respondent usually worked if they were employed during the reference period.
+- **UHRSWORK**: Reports the number of hours per week the respondent usually worked if they were employed during the reference period.
 
-INCWAGE: Represents each respondent's total pre-tax wage and salary income earned as an employee during the previous year.
+- **INCWAGE**: Represents each respondent's total pre-tax wage and salary income earned as an employee during the previous year.
 
-region: Constructed from the STATEFIP variable to represent geographic regions.
+- **education_level**: Derived from the EDUC variable, which indicates respondents' educational attainment as measured by the highest year of school or degree completed.
+It is grouped into five categories: Below High School (includes all levels below high school), High School (completed grade 12), Some College (completed 1-2 years of college), Bachelor’s Degree (completed 4 years of college), and Above Bachelor (more than 5 years of higher education).
+
+- **age**: Filtered from the AGE variable to include only respondents aged 18 to 65. This variable records the respondent’s age in years as of their last birthday.
+
+- **region**: Constructed from the STATEFIP variable to represent geographic regions.
 Valid state codes (1–56) were grouped into four traditional U.S.
@@
 
-education_level: Derived from the EDUC variable, which indicates respondents' educational attainment as measured by the highest year of school or degree completed.
-It is grouped into five categories: Below High School (includes all levels below high school), High School (completed grade 12), Some College (completed 1-2 years of college), Bachelor’s Degree (completed 4 years of college), and Above Bachelor (more than 5 years of higher education).age: Filtered from the AGE variable to include only respondents aged 18 to 65.
-This variable records the respondent’s age in years as of their last birthday.
+- **gender**: Based on the SEX variable, with original codes recoded as “Male” and “Female.”
 
-gender: Based on the SEX variable, with original codes recoded as “Male” and “Female.”
-
-race_group: Constructed from the RACE variable and grouped into four categories: White, Black, Asian, and Other.
+- **race_group**: Constructed from the RACE variable and grouped into four categories: White, Black, Asian, and Other.
 White and Black remain consistent with the original classifications.
@@
 plot1 <- ggplot(data, aes(x = UHRSWORK, y = INCWAGE)) +
-  geom_point(color = "blue", alpha = 0.6) +
+  geom_point(color = "steelblue", alpha = 0.6) +
   geom_smooth(method = "lm", formula = y ~ poly(x, 2), color = "red") +
@@
 plot2 <- ggplot(workhours_avg_income, aes(x = UHRSWORK, y = avg_income)) +
-  geom_line(color = "blue", size = 1) +
-  geom_point(color = "blue", size = 2) +
+  geom_line(color = "steelblue", size = 1) +
+  geom_point(color = "steelblue", size = 2) +
   labs(x = "Work Hours", y = "Average Income") +
@@
 plot3 <- ggplot(data, aes(x = age, y = INCWAGE)) +
-  geom_point(color = "orange", alpha = 0.6) +
+  geom_point(color = "darkorange", alpha = 0.6) +
   geom_smooth(method = "lm", formula = y ~ poly(x, 2), color = "red") +
@@
 plot4 <- ggplot(age_avg_income, aes(x = age, y = avg_income)) +
-  geom_line(color = "orange", size = 1) +
-  geom_point(color = "orange", size = 2) +
+  geom_line(color = "darkorange", size = 1) +
+  geom_point(color = "darkorange", size = 2) +
   labs(x = "Age", y = "Average Income") +
@@
   ) +
-  scale_fill_gradient(low = "white", high = "blue", name = "Proportion") +
+  scale_fill_gradient(low = "white", high = "steelblue", name = "Proportion") +
   scale_color_manual(values = c("black" = "black", "white" = "white"), guide = "none") +  # Define color mapping
@@
 
-The chart illustrates the proportion of different education levels across racial groups.
+@fig-Education-Race illustrates the proportion of different education levels across racial groups.
 The horizontal axis represents racial groups, including "Asian," "Black," "White," and "Other," while the vertical axis represents education levels, ranging from "Below High School" to "Above Bachelor." The intensity of blue indicates the proportion, with darker shades representing higher values.
@@
 
-The goal of our modelling strategy is twofold. Firstly,...
+The goal of our modelling strategy is to quantify the contributions of key factors—such as weekly working hours, age (and its quadratic term), education level, region, gender, and race—to variations in income (log-transformed).
+By estimating the relative impact of these predictors, the model provides insights into potential social and demographic disparities in income.
 
-Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
+Here we briefly describe the multiple linear regression model used to investigate these relationships.
+The model captures both linear and nonlinear effects (e.g., through the quadratic term for age), and all categorical variables are treated as factors to account for group-level differences.
+Background details, including diagnostics and model assumptions, are included in.
 
@@
 
-Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.
+This study use Multiple Linear Regression (MLR) to model the relationship between income and various predictors, implemented using R's `lm` function.
+The dataset is divided into training and testing sets using the `createDataPartition` function from the `caret` package, with 60% allocated for model training and parameter estimation and 40% for testing to evaluate predictive performance.
+
+Multiple regression models rely on several assumptions: linearity, meaning a linear relationship exists between independent and dependent variables; homoskedasticity, requiring constant error variance across predictors; independence of errors, meaning residuals are uncorrelated; normality, where residuals follow a normal distribution; and independence of independent variables, ensuring no significant multicollinearity.
+
+The final model is displayed below:
 
 \begin{align}
-y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
-\mu_i &= \alpha + \beta_i + \gamma_i\\
-\alpha &\sim \mbox{Normal}(0, 2.5) \\
-\beta &\sim \mbox{Normal}(0, 2.5) \\
-\gamma &\sim \mbox{Normal}(0, 2.5) \\
-\sigma &\sim \mbox{Exponential}(1)
+\log(\text{INCWAGE}_i) &= \beta_{0} + \beta_{1}\text{UHRSWORK}_i + \beta_{2}\text{region}_i \nonumber \\
+&\quad + \beta_{3}\text{education\_level}_i + \beta_{4}\text{age}_i + \beta_{5}\text{age}_i^2 \nonumber \\
+&\quad + \beta_{6}\text{gender}_i + \beta_{7}\text{race\_group}_i + \epsilon_i \label{eq:1}
 \end{align}
 
-We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
+- $\beta_{0}$ is the coefficient for the intercept.  
+- $\beta_{1}$ is the coefficient for the continuous variable $\text{UHRSWORK}_i$, which measures weekly hours worked.  
+- $\beta_{2}$ is the coefficient corresponding to the categorical variable $\text{region}_i$, which includes the levels Midwest, Northeast, South, and West. Midwest is the reference level.  
+- $\beta_{3}$ is the coefficient corresponding to the categorical variable $\text{education\_level}_i$, which includes the levels Above Bachelor, Bachelor, Below High School, High School, and Some College.
+Above Bachelor is the reference level.
+- $\beta_{4}$ and $\beta_{5}$ are the coefficients for the linear and quadratic terms of $\text{age}_i$, capturing the nonlinear relationship between age and income.  
+- $\beta_{6}$ is the coefficient for the binary variable $\text{gender}_i$, which includes the levels Male and Female. Female is the reference level.  
+- $\beta_{7}$ is the coefficient corresponding to the categorical variable $\text{race\_group}_i$, which includes the levels Asian, Black, Other, and White. Asian is the reference level.  
+- $\epsilon_i$ is the error term, capturing the deviation of the observed value from the predicted value due to unobserved factors.
 
@@
 
-We expect a positive relationship between the size of the wings and time spent aloft. In particular...
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
 
-We can use maths by including latex between dollar signs, for instance $\theta$.
+# Load the saved final model
+final_model <-
+  readRDS(file = here::here("models/final_model.rds"))
 
-# Results
+# Load the training and testing datasets
+train_data <- readRDS(file = here::here("models", "train_data.rds"))
+test_data <- readRDS(file = here::here("models", "test_data.rds"))
+```
 
-Our results are summarized in @tbl-modelresults.
+```{r}
+#| echo: false
+#| eval: true
+#| label: tbl-test
+#| tbl-cap: "Training and Testing Data Evaluation Results"
+#| warning: false
+#| results: asis
+
+library(caret)
+library(broom)
+library(knitr)
+library(kableExtra)
+
+# Predict on the training set
+train_data$predicted_log_INCWAGE <- predict(final_model, newdata = train_data)
+
+# Predict on the testing set
+test_data$predicted_log_INCWAGE <- predict(final_model, newdata = test_data)
+
+# Convert from log scale (exp) to original income values (optional)
+train_data$predicted_INCWAGE <- exp(train_data$predicted_log_INCWAGE)
+train_data$actual_INCWAGE <- exp(train_data$log_INCWAGE)
+test_data$predicted_INCWAGE <- exp(test_data$predicted_log_INCWAGE)
+test_data$actual_INCWAGE <- exp(test_data$log_INCWAGE)
+
+# Calculate error metrics for training and testing sets
+train_rmse <- sqrt(mean((train_data$predicted_log_INCWAGE - train_data$log_INCWAGE)^2))
+train_mae <- mean(abs(train_data$predicted_log_INCWAGE - train_data$log_INCWAGE))
+train_r_squared <- 1 - sum((train_data$predicted_log_INCWAGE - train_data$log_INCWAGE)^2) /
+  sum((train_data$log_INCWAGE - mean(train_data$log_INCWAGE))^2)
+
+test_rmse <- sqrt(mean((test_data$predicted_log_INCWAGE - test_data$log_INCWAGE)^2))
+test_mae <- mean(abs(test_data$predicted_log_INCWAGE - test_data$log_INCWAGE))
+test_r_squared <- 1 - sum((test_data$predicted_log_INCWAGE - test_data$log_INCWAGE)^2) /
+  sum((test_data$log_INCWAGE - mean(test_data$log_INCWAGE))^2)
+
+# Create a table summarizing training and testing performance
+performance_metrics <- data.frame(
+  Metric = c("RMSE", "MAE", "R²"),
+  Training = c(train_rmse, train_mae, train_r_squared),
+  Testing = c(test_rmse, test_mae, test_r_squared)
+)
+
+kable(
+  performance_metrics,
+  format = "latex",
+  digits = 3,
+  booktabs = TRUE
+) %>%
+  kable_styling(latex_options = "striped", position = "center")
+
+```
+
+The selection of variables and model structure was based on theory and data characteristics.
+Given the right-skewed nature of the income distribution in @fig-income-distribution, the dependent variable is log-transformed (`log(INCWAGE)`) to improve model fit.
+Furthermore, to capture the nonlinear relationship between age and income evident in @fig-combined-visualizations, a quadratic term for age (`age^2`) is included in the model.
+Categorical variables, such as region, education level, race group, and gender, are represented using dummy variables.
+Each category is compared to a reference category, allowing the regression model to capture and interpret group-level differences.
+
+During the model selection process, an initial model included an interaction term between education level and race group (education_level:race_group), based on the association observed in @fig-Education-Race, while other variables remained unchanged.
+This interaction term aimed to evaluate income differences across racial groups at the same education level.
+However, most interaction terms were not statistically significant and provided minimal improvement to model fit metrics, such as adjusted $R^2$ and AIC.
+To simplify the model and enhance interpretability, this interaction term was excluded from the final specification.
+
+### Model Results
+
+```{r}
+#| echo: false
+#| eval: true
+#| label: tbl-modelresults
+#| tbl-cap: "Estimated coefficients of the final model"
+#| warning: false
+#| results: asis
+
+# Prepare data: Extract model coefficients and confidence intervals
+model_summary <- data.frame(
+  Coefficients = coef(final_model),                
+  `Lower CI` = confint(final_model)[, 1],
+  `Upper CI` = confint(final_model)[, 2]
+)
+
+# Generate table using knitr::kable
+kable(
+  model_summary,
+  format = "latex",        
+  digits = 4              
+)
+
+```
 
@@
 
+```{r}
+
+```
+
 # Model details {#sec-model-details}
 
-## Posterior predictive check
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+#| label: fig-income-distribution
+#| fig-cap: "Distribution of Income: Original vs Log-transformed"
+#| fig-subcap: [
+#| "Income Distribution", 
+#| "Log-transformed Income Distribution"
+#| ]
+#| layout-ncol: 2
 
-In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...
+data <- data %>%
+  mutate(log_INCWAGE = log(INCWAGE))
 
-In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...
+# Original income distribution plot (without scientific notation)
+p1 <- ggplot(data, aes(x = INCWAGE)) +
+  geom_histogram(bins = 50, fill = "steelblue", color = "white", alpha = 0.8) +
+  scale_x_continuous(labels = scales::comma) +  # Use comma formatting for axis
+  labs(
+    x = "Income (INCWAGE)",
+    y = "Frequency"
+  ) +
+  theme_minimal()
+
+# Log-transformed income distribution plot
+p2 <- ggplot(data, aes(x = log_INCWAGE)) +
+  geom_histogram(bins = 50, fill = "darkorange", color = "white", alpha = 0.8) +
+  labs(
+    x = "Log of Income (log(INCWAGE))",
+    y = "Frequency"
+  ) +
+  theme_minimal()
+
+p1
+p2
+
+```
 

### 2024-11-25 scripts/06-model_data.R

@@ -11,27 +11,44 @@
 #### Workspace setup ####
 library(tidyverse)
 library(rstanarm)
+library(arrow)
+library(here)
+library(caret)
 
 #### Read data ####
-analysis_data <- read_csv("data/analysis_data/analysis_data.csv")
+data <- read_parquet(here("data", "01-analysis_data", "analysis_data.parquet"))
 
 ### Model data ####
-first_model <-
-  stan_glm(
-    formula = flying_time ~ length + width,
-    data = analysis_data,
-    family = gaussian(),
-    prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_aux = exponential(rate = 1, autoscale = TRUE),
-    seed = 853
-  )
 
+data <- data %>%
+  mutate(log_INCWAGE = log(INCWAGE))
+
+train_index <- createDataPartition(data$log_INCWAGE, p = 0.6, list = FALSE)  # 60% 训练集
+train_data <- data[train_index, ]
+test_data <- data[-train_index, ]
+
+
+# build a multiple linear regression model that includes a quadratic term for age
+final_model <- lm(
+  formula = log_INCWAGE ~ UHRSWORK + region + education_level + 
+    age + I(age^2) + gender + race_group, 
+  data = data
+)
+
+summary(final_model)
 
 #### Save model ####
 saveRDS(
-  first_model,
-  file = "models/first_model.rds"
+  final_model,
+  file = "models/final_model.rds"
 )
+
+saveRDS(train_data, file = "models/train_data.rds")
+saveRDS(test_data, file = "/test_data.rds")
+
+
+
+
+
 
 

### 2024-11-25 paper/paper.qmd
add model

@@
 library(kableExtra)
+library(car) 
+library(broom) 
 ```
@@
 The model captures both linear and nonlinear effects (e.g., through the quadratic term for age), and all categorical variables are treated as factors to account for group-level differences.
-Background details, including diagnostics and model assumptions, are included in.
 
@@
 Multiple regression models rely on several assumptions: linearity, meaning a linear relationship exists between independent and dependent variables; homoskedasticity, requiring constant error variance across predictors; independence of errors, meaning residuals are uncorrelated; normality, where residuals follow a normal distribution; and independence of independent variables, ensuring no significant multicollinearity.
+These assumptions are discussed in @sec-model-details and evaluated using the diagnostic plots shown in @fig-check.
 
@@
 
+The evaluation of the final model's performance was conducted on both training and testing datasets.
+As shown in tbl-test, the RMSE and MAE values are slightly higher for the testing data compared to the training data (0.786 vs.
+0.773 for RMSE, and 0.568 vs.
+0.543 for MAE), reflecting a modest decline in predictive accuracy when applied to unseen data.
+Similarly, the $R^2$ values are 0.488 for the training set and 0.468 for the testing set, indicating that the model captures a moderate proportion of the variability in the log-transformed income.
+Additionally, all variables included in the final model were statistically significant, with $p$-values below 0.05, further supporting the robustness of the selected predictors.
+Overall, the final model achieves a practical balance between complexity, interpretability, and predictive performance.
+
 ### Model Results
@@
 ```
+
+@tbl-modelresults presents the effects of weekly working hours, age, region, education level, gender, and racial group on log income, along with their 95% confidence intervals.
+The estimated coefficients indicate varying degrees of influence among these factors.
+
+Education level has the strongest impact on income.
+Compared to those with a high school education (the reference group), individuals with education beyond a bachelor’s degree have a coefficient of 1.084, showing a significant increase in income.
+Similarly, those with a bachelor’s degree and some college education have coefficients of 0.8713 and 0.5094, respectively, suggesting higher education correlates with higher income.
+
+Age and its quadratic term reveal a nonlinear relationship.
+The positive coefficient for age (0.0985) indicates income increases with age.
+However, the negative coefficient for the quadratic term (-0.0010) shows this growth slows over time, eventually declining.
+This pattern reflects typical career income trajectories, with rapid growth early on, stability in middle age, and potential decreases later in life.
+
+Gender also plays a significant role. Male income is significantly higher than female income, with a coefficient of 0.2632, even after controlling for other factors.
+
+The effect of region varies.
+Compared to the Midwest (reference group), incomes in the West and Northeast are higher, with coefficients of 0.1560 and 0.1287.
+The coefficient for the South is -0.0192, close to zero, indicating little difference.
+This may relate to regional economic conditions and policies.
+
+Weekly working hours show a modest effect on income, with a coefficient of 0.040.
+While longer hours increase income, the impact is relatively small, potentially due to wage structures and overtime policies.
+
+The effects of racial group are more complex.
+Compared to Asian individuals (reference group), income for Black and "Other" racial groups is notably lower, with coefficients of -0.1014 and -0.1290.
+The coefficient for White individuals is nearly zero (-0.0007), indicating no significant difference.
+These patterns reflect complexities in racial income disparities.
+
+In summary, education level has the largest influence on income, followed by age and gender.
+Region and racial group also play roles, while the effect of weekly working hours is smaller.
+These results provide a clear picture of how socioeconomic factors affect income and offer important context for further research and policymaking.
 
@@
 
+```{r}
+#| echo: false
+#| warning: false
+#| label: fig-check
+#| fig-cap: "Model Assumption Checks"
+#| fig-subcap: [
+#| "Residuals vs Fitted", "Normal Q-Q Plot", 
+#| "Scale-Location", "Residuals vs Leverage"
+#| ]
+#| layout-ncol: 2
+
+# Extract residuals and fitted values from the model
+model_data <- augment(final_model)
+
+# Check linearity assumption: Residuals vs Fitted Plot
+residuals_vs_fitted <- ggplot(model_data, aes(.fitted, .resid)) +
+  geom_point(alpha = 0.6) +
+  geom_smooth(method = "loess", se = FALSE, color = "red") +
+  labs(x = "Fitted values", y = "Residuals") +
+  theme_minimal()
+
+# Check normality of residuals: Normal Q-Q Plot
+qq_plot <- ggplot(model_data, aes(sample = .std.resid)) +
+  stat_qq() +
+  stat_qq_line(color = "red") +
+  labs(x = "Theoretical Quantiles", y = "Standardized Residuals") +
+  theme_minimal()
+
+# Check homoscedasticity: Scale-Location Plot
+scale_location <- ggplot(model_data, aes(.fitted, sqrt(abs(.std.resid)))) +
+  geom_point(alpha = 0.6) +
+  geom_smooth(method = "loess", se = FALSE, color = "red") +
+  labs(x = "Fitted values", y = "Sqrt(Standardized Residuals)") +
+  theme_minimal()
+
+# Check for outliers and leverage points: Residuals vs Leverage
+residuals_leverage <- ggplot(model_data, aes(.hat, .std.resid)) +
+  geom_point(alpha = 0.6) +
+  geom_smooth(method = "loess", se = FALSE, color = "red") +
+  labs(x = "Leverage", y = "Standardized Residuals") +
+  theme_minimal()
+
+residuals_vs_fitted
+qq_plot
+scale_location
+residuals_leverage
+
+```
+The model assumptions were checked using @fig-check.
+First, the Residuals vs Fitted plot tests the linearity assumption, showing that residuals are randomly scattered around the fitted values without any clear pattern, indicating that the linear relationship is valid.
+Second, the Normal Q-Q plot checks for the normality of residuals.
+Most points align closely with the diagonal line, suggesting the residuals are approximately normally distributed, with only slight deviations at the tails.
+Third, the Scale-Location plot evaluates homoskedasticity, and the red line remains mostly flat, suggesting that the variance of the residuals is fairly consistent.
+Finally, the Residuals vs Leverage plot identifies potential outliers or high-leverage points, indirectly testing the independence of errors and the independence of variables.
+Most points have low leverage, with only a few high-leverage points that may need attention.
+Overall, the assumptions are largely met, though there are some limitations that could be addressed with further refinements in the analysis.
+
 \newpage

### 2024-11-26 paper/paper.qmd
add discussion and introduction

@@
 ---
-title: "My title"
-subtitle: "My subtitle if needed"
+title: "Analysing the factors influencing the wages of the US workforce"
+subtitle: "Education as the Strongest Factor While Gender and Race Reflect Inequalities"
 author: 
-  - First author
-  - Another author
+  - Jianing Li
 thanks: "Code and data are available at: [https://github.com/RohanAlexander/starter_folder](https://github.com/RohanAlexander/starter_folder)."
@@
 date-format: long
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
+abstract: "This study uses a multiple linear regression model to analyze key factors influencing wages among the U.S.
+workforce, including education, age, gender, race, region, and hours worked.
+The results show that education is the strongest factor, with higher levels of education leading to significantly higher wages, while gender and race reveal notable disparities, with men and Asian workers earning the most.
+The study also finds a nonlinear relationship between age and wages, with region and hours worked playing smaller but significant roles.
+By identifying these drivers, the research provides evidence to inform policies aimed at reducing wage gaps and promoting fairer economic outcomes."
 format: pdf
@@
 
-Overview paragraph
+Income inequality is a persistent social and economic problem with significant implications for individual well-being and societal stability.
+Understanding how personal characteristics and external factors shape income distribution is essential to addressing its underlying causes.
+While many studies have explored income inequality, few have systematically analyzed the interaction of these factors.
+Additionally, existing research often lacks recent data that are both nationally representative and adaptable for diverse analyses.
+This study addresses these gaps using the 2023 sample from the IPUMS database, which allows users to customize variables and provides high-quality, comprehensive data.
+Leveraging this resource, the study identifies the main determinants of income inequality through a multiple linear regression model.
 
-Estimand paragraph
+The estimand in this study is income.
+Specifically, it quantifies the effects of education, age, gender, race, region, and hours worked on income.
+By measuring these effects, the study seeks to uncover the mechanisms driving income inequality and provide a framework for understanding how personal characteristics and external conditions interact to influence income distribution.
 
-Results paragraph
+The study finds that education is the strongest determinant of income.
+Higher levels of education lead to significantly higher earnings, with those with a college degree earning much more than other groups.
+The relationship between age and earnings is non-linear, with earnings peaking between the ages of 40 and 50 before declining, reflecting the impact of the occupational life cycle.
+In addition, gender and race show disparities, with men and Asians earning the highest average incomes, while women and certain minority groups earn less.
+Region and hours worked also affect earnings, but their effects are relatively small.
+Regional income differences reflect differences in economic development and policy environments, while hours worked show diminishing returns beyond a certain point.
+The findings are important because they reveal the role of education, gender, and race in driving income inequality and provide clear directions for policy interventions.
+They also provide empirical support for strategies to reduce income inequality, such as improving the allocation of educational resources, promoting gender equality, and increasing economic opportunities for minority groups.
 
-Why it matters paragraph
-
-Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....
+This paper provides a comprehensive analysis of income inequality by examining its key drivers and impacts.
+@sec-data describes the dataset in detail, highlights its main features, and uses visualizations to present the analysis results.
+@sec-model focuses on the construction of the multiple linear regression model, validates its effectiveness, and presents the results of the analysis.
+@sec-discussion analyzes the research findings, explores the reasons behind the observed patterns, and offers policy recommendations to address income inequality.
+It also identifies the study’s limitations and suggests directions for future research.
+@sec-appendix provides additional details on the data, model, and methods of survey and sampling, offering more comprehensive support for the study.
 
@@
 ## Data Overview
-Our data [@shelter]....
 
-We use the statistical programming language R [@citeR]....Our data [@shelter]....
+We used data from the IPUMS USA database [@ipums2024].
+All analyses were conducted in R [@citeR].
+Data simulation, testing, and cleaning were implemented using `tidyverse` [@tidyverse] and `testthat` [@testthat], with specific tasks performed using `here` [@here], `arrow` [@arrow], and `readr` [@readr].
+Data manipulation was carried out with `dplyr` [@dplyr] and `reshape2` [@reshape2], while data visualization was done using `ggplot2` [@ggplot2], adjusted with `scales` [@scales].
+Model training and performance evaluation were supported by `caret` [@caret], with model testing and analysis conducted using `broom` [@broom].
+Code formatting and table presentation were handled by `knitr` [@knitr] and `kableExtra` [@kableExtra], respectively, while regression diagnostics and additional analysis were performed with `car` [@car].
 
@@
 
-# Model
+# Model {#sec-model}
 
@@
 
-# Discussion
+# Discussion {#sec-discussion}
 
-## First discussion point {#sec-first-point}
+This paper uses a multiple linear regression model to analyze key factors that influence income, including education, age, gender, race, region, and hours worked.
+The study is based on data from the 2023 American Community Survey (ACS), cleaned and processed for visualization and analysis.
+The model quantifies the impact of these variables and explores their significance, showing how personal characteristics and external factors shape the income distribution.
+Education, age, gender, and race are treated as endogenous variables, reflecting individual attributes, while region and work hours are considered exogenous, capturing the influence of external conditions on income.
+Through the analysis of these variables, the study provides insights into the roots of income inequality and offers evidence to inform policymaking.
 
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
+## Education and Age: Core Drivers of Income
 
-## Second discussion point
+Education is a key factor influencing income, with a clear positive correlation between education level and earnings.
+Graduate degree holders earn significantly more, while a bachelor's degree also boosts income.
+However, those with only 1-2 years of college earn similar incomes to high school graduates, underscoring the importance of completing college.
+Higher education plays a crucial role in enhancing economic potential and securing high-paying jobs.
+Most income differences stem from variations in bachelor's, graduate, and doctoral education levels, while pre-college education has a smaller impact.
 
-Please don't use these as sub-heading labels - change them to be what your point actually is.
+Age also has a nonlinear impact on income.
+The model indicates that income peaks between ages 40 and 50 before gradually declining.
+The quadratic term captures this trend, reflecting how different stages of a career affect earnings.
+Early in their careers, individuals experience rapid income growth as they gain experience and promotions.
+In later stages, productivity declines, and income typically decreases as retirement approaches.
 
-## Third discussion point
+Education and age shape an individual's earning potential.
+These findings are valuable for career planning and offer guidance for policymakers.
+Individuals can enhance their competitiveness in the job market by pursuing a bachelor's degree or higher.
+Governments can increase education funding and provide financial aid to expand access to higher education.
+Additionally, support programs for those nearing retirement can help mitigate the impact of declining wages on their quality of life.
+
+## Gender and Race: Social Inequalities in Income Distribution
+
+Gender inequality in income distribution is evident.
+Studies show that men earn significantly more than women, even when controlling for education, age, and other factors.
+This disparity may stem from gender bias and occupational segregation.
+Women are often expected to take on more family responsibilities and face potential career interruptions due to childbirth [@gender].
+Additionally, men are overrepresented in many high-paying professions, contributing to lower average wages for women [@blau2013trends].
+
+Racial disparities, on the other hand, highlight another form of structural inequality.
+Research indicates that Asians earn the highest incomes, followed by Whites, while Black individuals and those from "other" racial groups earn the least.
+These differences can be attributed to historical and systemic factors, such as racial discrimination in the labor market and the underrepresentation of minorities in high-paying industries [@race].
+Educational disparities between racial groups, as shown in @fig-Education-Race, also contribute to these income gaps.
+
+Gender and race are exogenous factors, beyond individual control, and the wage disparities they cause reflect systemic inequality.
+Addressing the gender pay gap requires targeted policies, such as promoting pay transparency and increasing women's participation in high-paying industries.
+To reduce racial income disparities, policymakers can implement measures like providing education and employment support for minority groups, encouraging diversity hiring, and strictly enforcing anti-discrimination laws.
+Cultural change is also crucial; reducing biases against women and racial groups can help narrow wage gaps and promote fairness.
+
+## Region and Working Hours: External Environmental Impact on Income
+
+Region affects income.
+Studies show that incomes are higher in the Northeast and West and lowest in the Midwest.
+This may be due to differences in economic development and industry structures.
+Developed regions offer more high-paying jobs, while less developed areas have limited options.
+To reduce regional income gaps, the government can support industries in the Midwest and attract high-value sectors.
+Improving infrastructure and better allocating educational resources can also create more quality jobs in underdeveloped areas.
+
+Work hours are positively linked to income, but the returns diminish.
+Beyond 50 hours per week, income becomes more unstable.
+This suggests that working longer hours does not always lead to higher pay.
+Policies like overtime pay rules or limits on work hours may play a role.
+The government should introduce better labor laws, such as capping excessive work hours.
+Encouraging businesses to innovate and improve efficiency can reduce reliance on long hours.
+These measures can protect workers’ rights while promoting a fairer and healthier labor market.
 
@@
 
-Weaknesses and next steps should also be included.
+The data used in this study has certain limitations.
+It relies on survey responses, which can include self-reporting bias.
+For example, respondents may overestimate or underestimate their income or working hours.
+Additionally, the original data collection only classified gender as male or female.
+Non-binary and transgender groups were not included, leaving some populations unrepresented.
+Future research could incorporate more diverse datasets, such as administrative records or third-party data, to improve accuracy.
+Using more detailed classifications could also better reflect overlooked groups and enhance inclusivity.
+
+Some important variables were not included in the analysis, such as industry classification.
+While industry significantly affects income, its complex structure and lack of standardized categories made it difficult to analyze.
+This exclusion might have introduced bias into the results.
+To simplify the model, smaller racial groups, such as Japanese, Indigenous, and multiracial individuals, were categorized as “Other.” Although this made the analysis manageable, it overlooked finer details in racial differences.
+Future studies could refine these categories and include more detailed variables when sample sizes allow.
+
+The model also has limitations.
+With fewer variables, it cannot fully explain the complexities of income distribution.
+This study is based on U.S.
+data from 2023, so its findings may not apply to other countries or time periods.
+Future research could include more variables, such as industry, family background, and regional policies, to improve explanatory power.
+Combining time-series and cross-country data would also help study income trends and disparities on a broader s1cale.
 
@@
 
-# Appendix {-}
+# Appendix {-} {#sec-appendix}
 
 # Additional data details
-
-```{r}
-
-```
 
@@
 ```
+
 The model assumptions were checked using @fig-check.

### 2024-11-26 scripts/02-download_data.R

@@ -9,8 +9,8 @@
 
 
 #### Workspace setup ####
-library(opendatatoronto)
-library(tidyverse)
+
+
 # [...UPDATE THIS...]
 
 #### Download data ####

### 2024-11-26 scripts/06-model_data.R

@@ -10,7 +10,6 @@
 
 #### Workspace setup ####
 library(tidyverse)
-library(rstanarm)
 library(arrow)
 library(here)
 library(caret)
@@ -38,13 +37,21 @@ final_model <- lm(
 summary(final_model)
 
 #### Save model ####
+
 saveRDS(
   final_model,
-  file = "models/final_model.rds"
+  file = here("models", "final_model.rds")
 )
 
-saveRDS(train_data, file = "models/train_data.rds")
-saveRDS(test_data, file = "/test_data.rds")
+saveRDS(
+  train_data,
+  file = here("models", "train_data.rds")
+)
+
+saveRDS(
+  test_data,
+  file = here("models", "test_data.rds")
+)
 
 
 

### 2024-11-26 scripts/00-simulate_data.R
Add README

@@ -1,6 +1,6 @@
 #### Preamble ####
 # Purpose: Simulates a dataset to study the impact of several variables, 
-# including education, region, gender, and age, on wages. 
+# including education, working hour, region, gender, race and age, on wages. 
 # Author: Jianing Li
 # Date: 22 November 2024
 # Contact: lijianing.li@mail.utoronto.ca

### 2024-11-26 scripts/05-model_data.R

@@ -1,11 +1,13 @@
 #### Preamble ####
-# Purpose: Models... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Build a multiple linear regression model to explore the effect of each variable on wages
+# Author: Jianing Li
+# Date: 24 November 2024
+# Contact: lijianing.li@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: 
+#- The `tidyverse`， `caret`， `arrow` and `here` packages must be installed and loaded.
+#- 03-clean_data.R must have been run
+# Any other information needed? None
 
 
 #### Workspace setup ####

### 2024-11-26 scripts/02-download_data.R
add download data step

@@ -1,26 +1,0 @@
-#### Preamble ####
-# Purpose: Downloads and saves the data from [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
-# License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
-
-
-#### Workspace setup ####
-
-
-# [...UPDATE THIS...]
-
-#### Download data ####
-# [...ADD CODE HERE TO DOWNLOAD...]
-
-
-
-#### Save data ####
-# [...UPDATE THIS...]
-# change the_raw_data to whatever name you assigned when you downloaded it.
-write_csv(the_raw_data, "inputs/data/raw_data.csv") 
-
-         

### 2024-11-26 other/datasheet/datasheet_template.qmd
add Datasheet

---
title: "Datasheet for IPUMS USA Dataset"
author: 
  - Jianing Li
thanks: "Code and data are available at: https://github.com/JianingLi1225/Determinants-of-Wages.git"
date: today
date-format: long
abstract: "This datasheet provides a detailed description of a dataset compiled to analyze factors influencing wages among the U.S. workforce, based on data from the 2023 American Community Survey (ACS). The dataset includes over 1% of the U.S. population, highlight individuals aged 18 to 65 who are employed. It focuses on key variables such as education, age, gender, race, region, work hours, and income, each selected for their relevance to understanding the determinants of wages. The data has been preprocessed to ensure clarity and consistency, making it suitable for advanced statistical analysis and modeling."
format: pdf
number-sections: true
bibliography: references.bib
---

This datasheet extracts questions from @gebru2021datasheets to provide a structured framework for documenting datasets, ensuring transparency, accountability, and fostering responsible data use.

**Motivation**

1. *For what purpose was the dataset created? Was there a specific task in mind? Was there a specific gap that needed to be filled? Please provide a description.*
    - The dataset was created to analyze the factors influencing wages among the U.S. workforce. The aim of the study was to use a multiple linear regression model to investigate key factors such as education, age, gender, race, region, and hours worked on income. It addresses the gap in existing research, which often lacks recent, nationally representative data suitable for diverse analyses. The 2023 IPUMS USA sample fills this gap.
2. *Who created the dataset (for example, which team, research group) and on behalf of which entity (for example, company, institution, organization)?*
    - The dataset was created by the IPUMS team under the University of Minnesota. Variables selected by the author of the paper at the time of download.
3. *Who funded the creation of the dataset? If there is an associated grant, please provide the name of the grantor and the grant name and number.*
    - The creation of the dataset was funded by multiple organizations. Key contributors include the National Institutes of Health (NIH), specifically the Eunice Kennedy Shriver National Institute of Child Health and Human Development (NICHD) and the National Institute on Aging (NIA), which provided grants supporting various IPUMS projects. The National Science Foundation (NSF) has also been a significant funder of IPUMS initiatives. Additionally, the Bill and Melinda Gates Foundation has contributed funding to specific IPUMS projects. Alongside external funding, the University of Minnesota offers substantial institutional support to IPUMS, ensuring the sustainability and continued development of its datasets.
4. *Any other comments?*
    - None

**Composition**

1. *What do the instances that comprise the dataset represent (for example, documents, photos, people, countries)? Are there multiple types of instances (for example, movies, users, and ratings; people and interactions between them; nodes and edges)? Please provide a description.*
	- The instances represent individual members of the U.S. workforce. Each instance records attributes such as gender, age, race, education, region, hours worked, and income. The dataset does not contain multiple types of instances.
2. *How many instances are there in total (of each type, if appropriate)?*
	- The dataset is based on a 1% national sample from the 2023 American Community Survey (ACS)
3. *Does the dataset contain all possible instances or is it a sample (not necessarily random) of instances from a larger set? If the dataset is a sample, then what is the larger set? Is the sample representative of the larger set (for example, geographic coverage)? If so, please describe how this representativeness was validated/verified. If it is not representative of the larger set, please describe why not (for example, to cover a more diverse range of instances, because instances were withheld or unavailable).*
	- The dataset is a 1% random sample from a larger set of U.S. population data. Sampling weights and adjustments ensure that the sample is representative of the national population.
4. *What data does each instance consist of? "Raw" data (for example, unprocessed text or images) or features? In either case, please provide a description.*
	- Each instance contains preprocessed features, such as income (INCWAGE), weekly work hours (UHRSWORK), gender, age, race, education level, and region. The data has been cleaned and recoded for analysis.
5. *Is there a label or target associated with each instance? If so, please provide a description.*
	- Yes, the target variable for each instance is income (INCWAGE), which is used to analyze the effects of various factors on earnings.
6. *Is any information missing from individual instances? If so, please provide a description, explaining why this information is missing (for example, because it was unavailable). This does not include intentionally removed information, but might include, for example, redacted text.*
	- There is missing information. But the missing values are removed during the process of cleaning the data
7. *Are relationships between individual instances made explicit (for example, users' movie ratings, social network links)? If so, please describe how these relationships are made explicit.*
	- The instances in the dataset are primarily independent individual records, with no explicit relationships between them.
8. *Are there recommended data splits (for example, training, development/validation, testing)? If so, please provide a description of these splits, explaining the rationale behind them.*
	- The dataset is split into a training set (60%) and a testing set (40%). The training set is used for model training and parameter estimation, while the testing set evaluates the model's performance on unseen data.
9. *Are there any errors, sources of noise, or redundancies in the dataset? If so, please provide a description.*
	- Since the data is self-reported, there may be minor noise, such as inaccuracies in income or work hours. Additionally, smaller racial groups are categorized under "Other," which may lead to a loss of specific details.
10. *Is the dataset self-contained, or does it link to or otherwise rely on external resources (for example, websites, tweets, other datasets)? If it links to or relies on external resources, a) are there guarantees that they will exist, and remain constant, over time; b) are there official archival versions of the complete dataset (that is, including the external resources as they existed at the time the dataset was created); c) are there any restrictions (for example, licenses, fees) associated with any of the external resources that might apply to a dataset consumer? Please provide descriptions of all external resources and any restrictions associated with them, as well as links or other access points, as appropriate.*
	- The dataset is self-contained and does not rely on any external resources.
11. *Does the dataset contain data that might be considered confidential (for example, data that is protected by legal privilege or by doctor-patient confidentiality, data that includes the content of individuals' non-public communications)? If so, please provide a description.*
	- The dataset is anonymized and does not include any information that could directly identify individuals, so it does not contain confidential data.
12. *Does the dataset contain data that, if viewed directly, might be offensive, insulting, threatening, or might otherwise cause anxiety? If so, please describe why.*
	- No
13. *Does the dataset identify any sub-populations (for example, by age, gender)? If so, please describe how these subpopulations are identified and provide a description of their respective distributions within the dataset.*
	- Yes, the dataset identifies sub-populations based on gender, race, and region. For example, racial groups include White, Black, Asian, and "Other."
14. *Is it possible to identify individuals (that is, one or more natural persons), either directly or indirectly (that is, in combination with other data) from the dataset? If so, please describe how.*
	- No, it is not possible. The dataset has been strictly anonymized, making it impossible to directly or indirectly identify individuals.
15. *Does the dataset contain data that might be considered sensitive in any way (for example, data that reveals race or ethnic origins, sexual orientations, religious beliefs, political opinions or union memberships, or locations; financial or health data; biometric or genetic data; forms of government identification, such as social security numbers; criminal history)? If so, please provide a description.*
	- Yes, the dataset includes sensitive information such as race, gender, education, and income. These variables are necessary for studying socioeconomic phenomena.
16. *Any other comments?*
	- None

**Collection process**

1. *How was the data associated with each instance acquired? Was the data directly observable (for example, raw text, movie ratings), reported by subjects (for example, survey responses), or indirectly inferred/derived from other data (for example, part-of-speech tags, model-based guesses for age or language)? If the data was reported by subjects or indirectly inferred/derived from other data, was the data validated/verified? If so, please describe how.*
	- The data was collected through self-reported responses to the American Community Survey (ACS). To enhance accuracy, the ACS applies weighting methods to adjust for sampling and nonresponse biases and follows strict statistical methodologies in questionnaire design.
2. *What mechanisms or procedures were used to collect the data (for example, hardware apparatuses or sensors, manual human curation, software programs, software APIs)? How were these mechanisms or procedures validated?*
	- The data was collected using standardized survey questionnaires via online submissions, mail, and face-to-face interviews. The U.S. Census Bureau oversees the data collection process, ensuring that all procedures adhere to statistical standards.
3. *If the dataset is a sample from a larger set, what was the sampling strategy (for example, deterministic, probabilistic with specific sampling probabilities)?*
	- The dataset employs a 1% random sampling strategy from a larger U.S. population database. The sampling method ensures coverage of diverse geographic areas and population characteristics nationwide.
4. *Who was involved in the data collection process (for example, students, crowdworkers, contractors) and how were they compensated (for example, how much were crowdworkers paid)?*
	- The data was collected by staff and field surveyors from the U.S. Census Bureau. They are salaried employees compensated according to government pay scales.
5. *Over what timeframe was the data collected? Does this timeframe match the creation timeframe of the data associated with the instances (for example, recent crawl of old news articles)? If not, please describe the timeframe in which the data associated with the instances was created.*
	- The data was collected in 2023 and aligns with the time frame of the data instances.
6. *Were any ethical review processes conducted (for example, by an institutional review board)? If so, please provide a description of these review processes, including the outcomes, as well as a link or other access point to any supporting documentation.*
	- Data collection was overseen by the U.S. Census Bureau, adhering to all relevant ethical standards, including privacy protection and data anonymization. 
7. *Did you collect the data from the individuals in question directly, or obtain it via third parties or other sources (for example, websites)?*
	- The data was directly collected from individuals through their responses to the American Community Survey.
8. *Were the individuals in question notified about the data collection? If so, please describe (or show with screenshots or other information) how notice was provided, and provide a link or other access point to, or otherwise reproduce, the exact language of the notification itself.*
	- Yes, respondents are informed before the survey begins, with detailed explanations about the purpose of the survey and an overview of how the data will be used.
9. *Did the individuals in question consent to the collection and use of their data? If so, please describe (or show with screenshots or other information) how consent was requested and provided, and provide a link or other access point to, or otherwise reproduce, the exact language to which the individuals consented.*
	- Yes, participation in the survey is voluntary, and respondents provide consent by completing the questionnaire.
10. *If consent was obtained, were the consenting individuals provided with a mechanism to revoke their consent in the future or for certain uses? If so, please provide a description, as well as a link or other access point to the mechanism (if appropriate).*
	- Since the data is collected anonymously, respondents cannot withdraw the information they have submitted.
11. *Has an analysis of the potential impact of the dataset and its use on data subjects (for example, a data protection impact analysis) been conducted? If so, please provide a description of this analysis, including the outcomes, as well as a link or other access point to any supporting documentation.*
	- None
12. *Any other comments?*
	- None

**Preprocessing/cleaning/labeling**

1. *Was any preprocessing/cleaning/labeling of the data done (for example, discretization or bucketing, tokenization, part-of-speech tagging, SIFT feature extraction, removal of instances, processing of missing values)? If so, please provide a description. If not, you may skip the remaining questions in this section.*
	- The data cleaning process involved several steps to prepare the raw data for analysis. First, the dataset was filtered to include individuals aged 18 to 65, representing the working-age population, while records where the employment status was not "Employed" were excluded to focus on individuals with income. Categorical variables were recoded to enhance interpretability. For example, the education variable was grouped into five categories: "Below High School," "High School," "Some College," "Bachelor," and "Above Bachelor." The race variable was classified into "White," "Black," "Asian," and "Other." Regional data, based on state codes, was mapped into "Northeast," "Midwest," "South," and "West," and the gender variable was simplified to "Male" and "Female." For numerical variables, outliers in work hours were addressed by removing zero or extreme values to ensure validity. These steps ensured a clean and well-structured dataset, suitable for further analysis.
2. *Was the "raw" data saved in addition to the preprocessed/cleaned/labeled data (for example, to support unanticipated future uses)? If so, please provide a link or other access point to the "raw" data.*
	- The raw data is downloaded from the IPUMS website and saved locally on the computer.
3. *Is the software that was used to preprocess/clean/label the data available? If so, please provide a link or other access point.*
	- Yes, data cleaning was done using R and related packages, such as dplyr and tidyverse.

4. *Any other comments?*
	- None

**Uses**

1. *Has the dataset been used for any tasks already? If so, please provide a description.*
	- The dataset was used to analyze key factors influencing wages among the U.S. workforce, evaluating the effects of variables such as education, age, gender, race, region, and work hours using a multiple linear regression model.
2. *Is there a repository that links to any or all papers or systems that use the dataset? If so, please provide a link or other access point.*
	- A GitHub repository is provided, and the code and data can be accessed through https://github.com/JianingLi1225/Determinants-of-Wages.git
3. *What (other) tasks could the dataset be used for?*
	- The dataset can be used for studying income inequality, analyzing the impact of education, assessing racial and gender disparities, and evaluating regional economic policies.
4. *Is there anything about the composition of the dataset or the way it was collected and preprocessed/cleaned/labeled that might impact future uses? For example, is there anything that a dataset consumer might need to know to avoid uses that could result in unfair treatment of individuals or groups (for example, stereotyping, quality of service issues) or other risks or harms (for example, legal risks, financial harms)? If so, please provide a description. Is there anything a dataset consumer could do to mitigate these risks or harms?*
	- Since the data is self-reported, it may introduce minor biases. Additionally, smaller racial groups were categorized as "Other," which could limit detailed analysis. 
5. *Are there tasks for which the dataset should not be used? If so, please provide a description.*
	- The dataset is not suitable for tasks requiring personal identification or detailed tracking of individual changes, as the data is anonymized and lacks longitudinal information.
6. *Any other comments?*
	- None

**Distribution**

1. *Will the dataset be distributed to third parties outside of the entity (for example, company, institution, organization) on behalf of which the dataset was created? If so, please provide a description.*
	- Yes, IPUMS USA provides public access, and any qualified researcher can apply to use the data.
2. *How will the dataset be distributed (for example, tarball on website, API, GitHub)? Does the dataset have a digital object identifier (DOI)?*
	- The dataset is distributed via the IPUMS website, where users must create an account to download it.
3. *When will the dataset be distributed?*
	- The dataset is already available and updated annually with each new ACS release.
4. *Will the dataset be distributed under a copyright or other intellectual property (IP) license, and/or under applicable terms of use (ToU)? If so, please describe this license and/ or ToU, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms or ToU, as well as any fees associated with these restrictions.*
	- The dataset is distributed under IPUMS terms of use and is restricted to non-commercial research. Users must comply with the usage agreement.
5. *Have any third parties imposed IP-based or other restrictions on the data associated with the instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms, as well as any fees associated with these restrictions.*
	- No third-party restrictions, but users must agree to IPUMS terms and conditions.
6. *Do any export controls or other regulatory restrictions apply to the dataset or to individual instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any supporting documentation.*
	- No explicit mention of export controls.
7. *Any other comments?*
	- None

**Maintenance**

1. *Who will be supporting/hosting/maintaining the dataset?*
	- The dataset is supported, hosted, and maintained by IPUMS, which is part of the Institute for Social Research and Data Innovation at the University of Minnesota.
2. *How can the owner/curator/manager of the dataset be contacted (for example, email address)?*
	- IPUMS can be contacted via email at ipums@umn.edu for feedback, questions, or concerns.
3. *Is there an erratum? If so, please provide a link or other access point.*
	- IPUMS maintains a revisions page where updates and corrections to datasets are documented. For example, changes made to Version 7.4 before archiving Version 7.5 are detailed on the current revisions page. https://international.ipums.org/international-action/revisions
4. *Will the dataset be updated (for example, to correct labeling errors, add new instances, delete instances)? If so, please describe how often, by whom, and how updates will be communicated to dataset consumers (for example, mailing list, GitHub)?*
	- IPUMS datasets are continually revised to incorporate updates, corrections, or improvements. These changes are documented on the respective project's revisions page, such as for the IPUMS USA project.
5. *If the dataset relates to people, are there applicable limits on the retention of the data associated with the instances (for example, were the individuals in question told that their data would be retained for a fixed period of time and then deleted)? If so, please describe these limits and explain how they will be enforced.*
	- IPUMS adheres to ethical standards and legal requirements concerning data retention. Its digital preservation framework emphasizes retaining data and metadata to meet archival requirements of funding agencies while ensuring data quality.
6. *Will older versions of the dataset continue to be supported/hosted/maintained? If so, please describe how. If not, please describe how its obsolescence will be communicated to dataset consumers.*
	- Yes, IPUMS archives annual snapshots of its datasets and documents changes between versions, ensuring that older versions remain accessible for research and reference purposes.
7. *If others want to extend/augment/build on/contribute to the dataset, is there a mechanism for them to do so? If so, please provide a description. Will these contributions be validated/verified? If so, please describe how. If not, why not? Is there a process for communicating/distributing these contributions to dataset consumers? If so, please provide a description.*
	- While IPUMS does not have a formal mechanism for external contributions to its datasets, it collaborates with national statistical offices and other organizations to collect and harmonize data.
8. *Any other comments?*
	- None

\newpage

# References

### 2024-11-26 paper/paper.qmd
add Datasheet

@@
   - Jianing Li
-thanks: "Code and data are available at: [https://github.com/RohanAlexander/starter_folder](https://github.com/RohanAlexander/starter_folder)."
+thanks: "Code and data are available at: https://github.com/JianingLi1225/Determinants-of-Wages.git"
 date: today
@@
 It also identifies the study’s limitations and suggests directions for future research.
-@sec-appendix provides additional details on the data, model, and methods of survey and sampling, offering more comprehensive support for the study.
+@sec-appendix provides additional details on the model, and methods of survey and sampling, offering more comprehensive support for the study.
 
@@
 
-# Additional data details
-
-# Model details {#sec-model-details}
+## Model details {#sec-model-details}
 

### 2024-11-26 scripts/05-model_data.R

@@ -24,7 +24,7 @@ data <- read_parquet(here("data", "01-analysis_data", "analysis_data.parquet"))
 data <- data %>%
   mutate(log_INCWAGE = log(INCWAGE))
 
-train_index <- createDataPartition(data$log_INCWAGE, p = 0.6, list = FALSE)  # 60% 训练集
+train_index <- createDataPartition(data$log_INCWAGE, p = 0.6, list = FALSE)  
 train_data <- data[train_index, ]
 test_data <- data[-train_index, ]
 

### 2024-11-26 paper/paper.qmd
add weakness

@@
 This exclusion might have introduced bias into the results.
-To simplify the model, smaller racial groups, such as Japanese, Indigenous, and multiracial individuals, were categorized as “Other.” Although this made the analysis manageable, it overlooked finer details in racial differences.
+To simplify the model, smaller racial groups, such as Indigenous, and multiracial individuals, were categorized as “Other.” Although this made the analysis manageable, it overlooked finer details in racial differences.
 Future studies could refine these categories and include more detailed variables when sample sizes allow.

### 2024-11-26 paper/paper.qmd
fix reference

@@
 library(broom) 
+library(caret)
+
 ```
@@
 
-library(caret)
-library(broom)
-library(knitr)
-library(kableExtra)
-
 # Predict on the training set
@@
 
-# Appendix {-} {#sec-appendix}
+# Appendix  {#sec-appendix}
 

### 2024-11-28 paper/paper.qmd
Add appendix

@@
 Nonetheless, the use of sampling weights and robust design make the dataset a reliable foundation for studying wage income and its associated factors.
-A detailed explanation of the survey methodology and sampling design is provided in the appendix.
 
@@
 Future research could include more variables, such as industry, family background, and regional policies, to improve explanatory power.
-Combining time-series and cross-country data would also help study income trends and disparities on a broader s1cale.
+Combining time-series and cross-country data would also help study income trends and disparities on a broader scale.
 
@@
 
+## Idealized Methodology
+
+### Overview
+
+This survey is designed to collect data on factors influencing wages across the United States.
+The focus is on education, gender, race, region, and working hours.
+The budget is $100,000, covering sampling design, respondent recruitment, data collection, and quality control.
+The goal is to ensure high-quality, representative data.
+The results will help develop fair economic policies, reduce wage gaps, and promote equity.
+
+### Sampling Approach
+
+The survey uses stratified random sampling to ensure a comprehensive and representative sample.
+Respondents are grouped by gender, education, race, and geographic region.
+Gender categories include male, female, and non-binary/transgender.
+Education is grouped into high school or below, some college, and bachelor’s degree or above.
+Race categories are White, Black, Asian, and Other (e.g., mixed or minority groups).
+Geographic regions include Northeast, Midwest, South, and West.
+Exact age is collected as a continuous variable.
+The total sample size is 5,000, with proportional representation for each group.
+This method ensures diversity and provides a strong foundation for analysis.
+
+### Recruitment
+
+The survey combines online and offline recruitment methods to reach diverse participants.
+Online ads are placed on platforms like Google, Facebook, and LinkedIn.
+These ads target individuals based on occupation and location.
+The content emphasizes privacy and the importance of participation, with messages like, “Take 3 minutes to help uncover the key drivers of wage differences.” This approach is effective for urban residents, younger people, and those with higher education levels.
+
+For groups less likely to respond to online ads, such as older adults and rural residents, Random Digit Dialing (RDD) is used.
+Trained interviewers make calls across all states and record responses directly.
+This ensures inclusivity by reaching populations underrepresented in online recruitment.
+
+To encourage participation, respondents are offered incentives.
+For every 100 surveys completed, one respondent is randomly chosen to receive a $5-$10 gift card.
+This incentive is explained at the start of the survey to motivate completion.
+
+### Data Collection and Survey Design
+
+Data is collected primarily through Google Forms.
+Respondents can complete the survey on a computer or mobile device.
+Each Google account is limited to one submission to prevent duplicates.
+Responses from RDD participants are manually entered into the same system for consistency.
+
+The survey includes attention check questions to ensure valid responses.
+For example, a question may ask respondents to “Select ‘Other’ to confirm you are paying attention.” Responses failing these checks are marked invalid.
+This improves data quality without adding extra burden to respondents.
+
+### Data Validation and Quality Control
+
+The survey implements several measures to ensure data accuracy and representativeness.
+Post-stratification weighting adjusts the sample to align with U.S.
+population characteristics.
+Logical checks identify and remove inconsistent responses, such as reports of zero work hours with high income.
+Duplicate submissions are identified and removed using account and IP address information.
+Missing data for open-ended questions, such as income, is handled using imputation.
+These steps improve the reliability and quality of the data.
+
+### Multi-Wave Data Collection and Aggregation
+
+To capture changes in the labor market, the survey is conducted in three waves, two months apart.
+This approach tracks trends and seasonal variations.
+After each round, data is reviewed and aggregated using weighted averages.
+This reduces random fluctuations and provides more accurate results for analysis.
+
+### Budget Allocation
+
+The budget is allocated based on the importance of each task.
+Approximately $40,000 is spent on online ads to reach a wide audience.
+Another $20,000 is used for RDD to connect with groups less accessible online, such as older adults and rural residents.
+An additional $15,000 is set aside for incentives to increase completion rates.
+The remaining $25,000 is allocated for data cleaning, validation, and multi-wave integration.
+This ensures efficient resource use and reliable survey outcomes.
+
+## Idealized survey
+
+The idealized survey questionnaire can be accessed via the following link： https://forms.gle/Aw9cjdqL9tB9bNHr8
+
+### Survey Questions
+
+Thank you for participating in this survey!
+This study aims to understand the factors influencing wages in the United States.
+Your responses will remain strictly confidential and will only be used for research purposes.
+The survey takes approximately 3-5 minutes to complete.
+
+As a token of appreciation, participants who complete the survey will have the chance to win a $5-$10 gift card. One respondent out of every 100 will be randomly selected to receive this reward.
+
+For Questions or Concerns, Please Contact:
+
+Jianing Li  
+Email: lijianing.li@mail.utoronto.ca
+
+1. What is your age?  
+   - Open-ended (numeric input)
+
+2. What is your gender?  
+   - Male  
+   - Female  
+   - Non-binary/Transgender  
+   - Prefer not to say  
+   - Other
+
+3. What is your highest level of education?  
+   - High school or below  
+   - Some college  
+   - Bachelor’s degree  
+   - Master’s degree or above
+
+4. What is your race or ethnicity?  
+   - White  
+   - Black  
+   - Asian  
+   - Other (please specify): ________
+
+5. Which U.S. state do you currently reside in?  
+   - Dropdown list of all 50 states
+
+6. To confirm you are paying attention, please select "Other" as your answer.  
+   - Male  
+   - Female  
+   - Other  
+   - Prefer not to say
+
+7. What is your primary employment status?  
+   - Full-time  
+   - Part-time  
+   - Self-employed  
+   - Unemployed  
+   - Retired  
+   - Student
+
+8. How many hours do you typically work per week?  
+   - Open-ended (numeric input)
+
+9. What is your annual pre-tax wage or salary income?  
+   - Open-ended (numeric input)
+
+10. How many years have you been working in your current field or profession?  
+   - Open-ended (numeric input)
+
+11. What is the primary reason you chose your current occupation?  
+   - Interest/passion  
+   - Financial stability  
+   - Family influence  
+   - Availability of jobs in the area  
+   - Other (please specify): ________
+
+If you would like to participate in the reward draw, please provide your email address.  
+   - Open-ended
+
+Thank You!  
+Thank you for your time and participation in this survey. Your responses are valuable to our research.
+
+Good luck with the reward draw!
+
 \newpage

### 2024-11-28 paper/paper.qmd
fix typo
add appendix

@@
 title: "Analysing the factors influencing the wages of the US workforce"
-subtitle: "Education as the Strongest Factor While Gender and Race Reflect Inequalities"
+subtitle: "Higher Education and Male Gender as Key Determinants of Wage Increases Controlling for Region and Race"
 author: 
@@
 library(caret)
+library(modelsummary)
 
@@
 Model training and performance evaluation were supported by `caret` [@caret], with model testing and analysis conducted using `broom` [@broom].
+Model summaries were generated using the `modelsummary` package [@modelsummary].
 Code formatting and table presentation were handled by `knitr` [@knitr] and `kableExtra` [@kableExtra], respectively, while regression diagnostics and additional analysis were performed with `car` [@car].
@@
 These variables cover geographic location, demographic characteristics, education, and employment conditions, providing a broad basis for examining factors influencing wage income.
+The cleaned dataset is presented in @tbl-datasummary, along with detailed descriptions of variable definitions and construction methods, which are thoroughly explained in @sec-data.
 
@@
 
-This study utilizes sample data from the 2023 ACS.
+This study uses sample data from the 2023 American Community Survey (ACS).
 The dataset is based on a 1% random national sample, ensuring broad representativeness.
-It includes individuals living in private households as well as those in group quarters, such as student dormitories and care facilities.
-Weights were applied to the data to account for potential biases introduced by the sampling design and nonresponse.
-
-Data collection for the ACS involved structured questionnaires covering a wide range of topics, such as demographics, education, employment, and income.
-For example, UHRSWORK (usual hours worked per week) captures labor input based on respondents’ reports of their average weekly work hours.
-The smallest geographic unit in the dataset is the Public Use Microdata Area (PUMA), which contains at least 100,000 residents and falls within state boundaries.
-This structure protects respondent privacy while supporting regional analysis.
-
-While the ACS methodology ensures data reliability and extensive coverage, variables like income and work hours rely on self-reported responses, which may introduce minor inaccuracies.
-Nonetheless, the use of sampling weights and robust design make the dataset a reliable foundation for studying wage income and its associated factors.
-
-## Dataset Description
-
-```{r}
-#| echo: false
-#| eval: true
-#| label: tbl-datasummary
-#| tbl-cap: "Cleaned Data Overview"
-#| warning: false
-#| results: asis
-
-# Ensure all categories of each categorical variable are displayed
-complete_data <- data %>%
-  group_by(education_level) %>%      
-  summarise_all(~ first(.)) %>%       # Ensure at least one row is displayed for each category
-  select(INCWAGE, UHRSWORK, everything())
-
-kable(
-  complete_data, 
-  align = 'c',
-  format = "latex"
-) %>%
-  kable_styling(
-    latex_options = "scale_down",  
-    full_width = FALSE
-  ) %>%
-  column_spec(1, width = "3cm")          
-```
-
-The sample of the cleaned dataset is shown in @tbl-datasummary, including the following variables.
-UHRSWORK and INCWAGE are retained from the raw data, while the other variables are newly constructed.
-During the data cleaning process, only individuals aged 18-65 and those with an employment status of "Employed" were included.
-This ensures the dataset focuses on studying the wage determinants of the employed labor force.
-Below is a description of each variable and how it was constructed:
-
-- **UHRSWORK**: Reports the number of hours per week the respondent usually worked if they were employed during the reference period.
-
-- **INCWAGE**: Represents each respondent's total pre-tax wage and salary income earned as an employee during the previous year.
-
-- **education_level**: Derived from the EDUC variable, which indicates respondents' educational attainment as measured by the highest year of school or degree completed.
-It is grouped into five categories: Below High School (includes all levels below high school), High School (completed grade 12), Some College (completed 1-2 years of college), Bachelor’s Degree (completed 4 years of college), and Above Bachelor (more than 5 years of higher education).
-
-- **age**: Filtered from the AGE variable to include only respondents aged 18 to 65. This variable records the respondent’s age in years as of their last birthday.
-
-- **region**: Constructed from the STATEFIP variable to represent geographic regions.
-Valid state codes (1–56) were grouped into four traditional U.S.
-regions: Northeast, Midwest, South, and West.
-Each state was assigned to its corresponding region for further regional analysis.
-
-- **gender**: Based on the SEX variable, with original codes recoded as “Male” and “Female.”
-
-- **race_group**: Constructed from the RACE variable and grouped into four categories: White, Black, Asian, and Other.
-White and Black remain consistent with the original classifications.
-Asian includes Chinese, Japanese, and Other Asian categories, while Other includes all other races and mixed-race groups.
+The sample includes individuals living in private households and those in group quarters, such as student dormitories, correctional facilities, and care homes.
+Weights are applied to the data to reduce biases caused by the sampling design and nonresponse.
+This improves the generalizability of the results.
+The smallest geographic unit in the dataset is the Public Use Microdata Area (PUMA).
+Each PUMA includes at least 100,000 residents and is contained within state boundaries.
+This design anonymizes respondents’ locations while enabling regional analysis.
+
+The ACS collects data using multiple methods.
+These include mailed questionnaires, online submissions, and in-person or telephone interviews.
+The surveys cover a wide range of topics, such as demographics, education, employment, and income.
+For example, wage income is recorded by asking respondents to report their total pre-tax earnings from wages, salaries, tips, and other sources over the past 12 months.
+Similarly, the variable UHRSWORK (usual hours worked per week) measures labor input based on respondents’ self-reported average weekly work hours during the survey.
+
+Despite its robust sampling methodology and extensive coverage, the ACS data has some inherent limitations.
+Variables such as income and work hours rely heavily on self-reported responses.
+This introduces potential inaccuracies due to recall bias, rounding errors, or deliberate misreporting.
+For individuals with irregular or multiple income sources, questions about wages or hours worked may be interpreted differently, further increasing measurement errors.
+
+To address these issues, the ACS uses data validation checks and imputation techniques to handle incomplete or inconsistent responses.
+However, these methods cannot fully eliminate the risk of bias or error.
+Even with these limitations, the ACS dataset remains a reliable and widely used resource for studying wage income and related factors.
+Its combination of sampling weights, large sample size, and rigorous design framework supports robust analysis.
 
@@
 
-Chart (a) is a scatter plot of work hours versus income, with each point representing an individual.
-The red quadratic regression line indicates a positive correlation between income and work hours, where longer hours generally lead to higher income.
-However, the regression line primarily reflects a linear trend, suggesting a stable relationship between income and work hours without significant signs of deceleration.
-Additionally, income variability increases with longer work hours, particularly beyond 50 hours per week, where fluctuations become more pronounced.
-Chart (b) illustrates the relationship between work hours and average income.
-While income generally increases with work hours, especially between 20 and 50 hours per week, it becomes unstable beyond 50 hours.
-
-Chart (c) is a scatter plot of age versus income, with the red quadratic regression line showing that income increases with age before eventually declining.
-Chart (d) depicts the trend of average income by age.
-The line chart shows that average income steadily rises with age, peaking in the late 40s and early 50s, and gradually declines as individuals approach retirement age.
+Chart (a) visualizes the relationship between work hours and income, where the red quadratic regression line indicates a positive correlation.
+Longer work hours generally correspond to higher income, with the regression line reflecting a stable linear trend rather than noticeable deceleration.
+Income variability increases with longer work hours, particularly beyond 50 hours per week, where fluctuations become more pronounced.
+Chart (b) represents the trend of average income by work hours.
+Income shows a steady increase with work hours, especially between 20 and 50 hours per week, but becomes less consistent beyond 50 hours.
+
+Chart (c) illustrates the relationship between age and income, with the red quadratic regression line highlighting a pattern where income rises with age before gradually declining.
+Chart (d) presents average income by age as a line chart.
+It shows that average income steadily increases with age, peaking in the late 40s and early 50s, followed by a gradual decline as individuals approach retirement age.
 
@@
 
-@fig-Education-Race illustrates the proportion of different education levels across racial groups.
-The horizontal axis represents racial groups, including "Asian," "Black," "White," and "Other," while the vertical axis represents education levels, ranging from "Below High School" to "Above Bachelor." The intensity of blue indicates the proportion, with darker shades representing higher values.
+@fig-Education-Race illustrates the proportion of different education levels across racial groups. The intensity of blue indicates the proportion, with darker shades representing higher values.
 
@@
 
-Overall, the distribution of education levels varies across racial groups.
-Asians have a higher proportion of highly educated individuals compared to other groups, while "High School" education is more prevalent among Black and "Other" groups.
-This reflects the complex relationship between race and educational background.
+Educational attainment shows significant variation across racial groups, reflecting the complex relationship between race and educational background.
+Asians have the highest levels of education, with a strong representation in bachelor's and advanced degrees.
+In contrast, Black and White groups display a broader distribution, with high school education being more common.
+The "Other" category, encompassing smaller racial groups and mixed-race individuals, stands out with a concentration at the high school level and lower representation in higher education, suggesting potential socio-economic and cultural disparities.
 
@@
 
-## Model set-up
+## Model Set-up
 
@@
 
-### Model justification
+### Model Justification
 
@@
 #| eval: true
-#| label: tbl-modelresults
-#| tbl-cap: "Estimated coefficients of the final model"
-#| warning: false
-#| results: asis
-
-# Prepare data: Extract model coefficients and confidence intervals
-model_summary <- data.frame(
-  Coefficients = coef(final_model),                
-  `Lower CI` = confint(final_model)[, 1],
-  `Upper CI` = confint(final_model)[, 2]
+#| label: fig-coefficients
+#| fig-cap: "Predictor Coefficients with Confidence Intervals"
+#| warning: false
+
+# Extract data from the model
+model_data <- modelplot(
+  final_model,
+  conf_level = 0.95,  
+  draw = FALSE        
 )
 
-# Generate table using knitr::kable
-kable(
-  model_summary,
-  format = "latex",        
-  digits = 4              
+# Filter out the Intercept
+filtered_data <- model_data[model_data$term != "(Intercept)", ]
+
+# Create a coefficient plot without Intercept
+ggplot(filtered_data, aes(x = estimate, y = term)) +
+  geom_point(size = 2, color = "blue") +              
+  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), 
+                 height = 0.2, color = "darkgray") +
+  theme_minimal() +                                   
+  labs(
+    x = "Coefficient Estimate",
+    y = "Predictors"
   )
@@
 
-@tbl-modelresults presents the effects of weekly working hours, age, region, education level, gender, and racial group on log income, along with their 95% confidence intervals.
-The estimated coefficients indicate varying degrees of influence among these factors.
-
-Education level has the strongest impact on income.
-Compared to those with a high school education (the reference group), individuals with education beyond a bachelor’s degree have a coefficient of 1.084, showing a significant increase in income.
-Similarly, those with a bachelor’s degree and some college education have coefficients of 0.8713 and 0.5094, respectively, suggesting higher education correlates with higher income.
-
-Age and its quadratic term reveal a nonlinear relationship.
-The positive coefficient for age (0.0985) indicates income increases with age.
-However, the negative coefficient for the quadratic term (-0.0010) shows this growth slows over time, eventually declining.
-This pattern reflects typical career income trajectories, with rapid growth early on, stability in middle age, and potential decreases later in life.
-
-Gender also plays a significant role. Male income is significantly higher than female income, with a coefficient of 0.2632, even after controlling for other factors.
-
-The effect of region varies.
-Compared to the Midwest (reference group), incomes in the West and Northeast are higher, with coefficients of 0.1560 and 0.1287.
-The coefficient for the South is -0.0192, close to zero, indicating little difference.
-This may relate to regional economic conditions and policies.
-
-Weekly working hours show a modest effect on income, with a coefficient of 0.040.
-While longer hours increase income, the impact is relatively small, potentially due to wage structures and overtime policies.
-
-The effects of racial group are more complex.
-Compared to Asian individuals (reference group), income for Black and "Other" racial groups is notably lower, with coefficients of -0.1014 and -0.1290.
-The coefficient for White individuals is nearly zero (-0.0007), indicating no significant difference.
-These patterns reflect complexities in racial income disparities.
-
-In summary, education level has the largest influence on income, followed by age and gender.
-Region and racial group also play roles, while the effect of weekly working hours is smaller.
-These results provide a clear picture of how socioeconomic factors affect income and offer important context for further research and policymaking.
+@fig-coefficients shows the effects of weekly working hours, age, region, education level, gender, and racial group on income.
+The coefficient plot also displays the 95% confidence intervals for each coefficient, with detailed numerical values presented in @tbl-modelresults in @sec-model-details.
+
+Education level is the most important factor.
+Individuals with lower education, such as "Below High School," have negative coefficients, while those with higher education, such as a "Bachelor’s" degree, show positive coefficients.
+This highlights the significant economic benefits of higher education.
+However, the confidence intervals for education vary.
+For example, "Below High School" has a wider interval, indicating greater uncertainty in the estimate.
+
+Age also has a strong impact on income.
+The positive coefficient for age shows that income increases as people get older.
+However, the negative coefficient for age squared suggests that this growth slows over time and may decline later in life.
+The confidence interval for age is narrow, indicating a precise estimate.
+Gender differences are also evident.
+Male individuals have significantly higher income, as shown by the positive coefficient for gender.
+The narrow confidence interval for gender confirms the reliability of this effect.
+
+Racial and ethnic factors present more complex dynamics.
+For example, "Black" and "Other" groups have negative coefficients and wider confidence intervals, reflecting greater uncertainty in their effects.
+In contrast, "White" individuals have coefficients close to zero and narrow intervals, showing minimal differences from the reference group.
+Regional effects are also notable.
+The "West" has positive coefficients compared to the reference region, while other regions show smaller or slightly negative effects.
+Some regional variables have confidence intervals that include zero, suggesting these effects may not be statistically significant.
+
+Finally, the effect of weekly working hours on income is relatively small but still positive.
+Its narrow confidence interval indicates a stable and reliable estimate.
+Overall, these results provide a clear understanding of the socioeconomic factors influencing income and offer valuable insights for future research and policy discussions.
 
@@
 
-## Weaknesses and next steps
+## Weaknesses and Next Steps
 
@@
 
-## Model details {#sec-model-details}
+## Dataset Description {#sec-data}
+
+```{r}
+#| echo: false
+#| eval: true
+#| label: tbl-datasummary
+#| tbl-cap: "Cleaned Data Overview"
+#| warning: false
+#| results: asis
+
+# Ensure all categories of each categorical variable are displayed
+complete_data <- data %>%
+  group_by(education_level) %>%      
+  summarise_all(~ first(.)) %>%       # Ensure at least one row is displayed for each category
+  select(INCWAGE, UHRSWORK, everything())
+
+kable(
+  complete_data, 
+  align = 'c',
+  format = "latex"
+) %>%
+  kable_styling(
+    latex_options = "scale_down",  
+    full_width = FALSE
+  ) %>%
+  column_spec(1, width = "3cm")  
+```
+
+The sample of the cleaned dataset is shown in @tbl-datasummary, including the following variables.
+UHRSWORK and INCWAGE are retained from the raw data, while the other variables are newly constructed.
+During the data cleaning process, only individuals aged 18-65 and those with an employment status of "Employed" were included.
+This ensures the dataset focuses on studying the wage determinants of the employed labor force.
+Below is a description of each variable and how it was constructed:
+
+- **UHRSWORK**: Reports the number of hours per week the respondent usually worked if they were employed during the reference period.
+
+- **INCWAGE**: Represents each respondent's total pre-tax wage and salary income earned as an employee during the previous year.
+
+- **education_level**: Derived from the EDUC variable, which indicates respondents' educational attainment as measured by the highest year of school or degree completed.
+It is grouped into five categories: Below High School (includes all levels below high school), High School (completed grade 12), Some College (completed 1-2 years of college), Bachelor’s Degree (completed 4 years of college), and Above Bachelor (more than 5 years of higher education).
+
+- **age**: Filtered from the AGE variable to include only respondents aged 18 to 65. This variable records the respondent’s age in years as of their last birthday.
+
+- **region**: Constructed from the STATEFIP variable to represent geographic regions.
+Valid state codes (1–56) were grouped into four traditional U.S.
+regions: Northeast, Midwest, South, and West.
+Each state was assigned to its corresponding region for further regional analysis.
+
+- **gender**: Based on the SEX variable, with original codes recoded as “Male” and “Female.”
+
+- **race_group**: Constructed from the RACE variable and grouped into four categories: White, Black, Asian, and Other.
+White and Black remain consistent with the original classifications.
+Asian includes Chinese, Japanese, and Other Asian categories, while Other includes all other races and mixed-race groups.
+
+## Model Details {#sec-model-details}
 
@@
 
+```{r}
+#| echo: false
+#| eval: true
+#| label: tbl-modelresults
+#| tbl-cap: "Estimated coefficients of the final model"
+#| warning: false
+#| results: asis
+
+# Prepare data: Extract model coefficients and confidence intervals
+model_summary <- data.frame(
+  Coefficients = coef(final_model),                
+  `Lower CI` = confint(final_model)[, 1],
+  `Upper CI` = confint(final_model)[, 2]
+)
+
+# Generate table using knitr::kable
+kable(
+  model_summary,
+  format = "latex",        
+  digits = 4              
+)
+```
+
 ## Idealized Methodology

### 2024-11-28 paper/paper.qmd
Auto stash before merge of "main" and "origin/main"
Final Commit

@@
 number-sections: true
+toc: true
 bibliography: references.bib
@@
 library(modelsummary)
+library(styler)
 
@@
 
-# Introduction
+# Introduction {#sec-Introduction}
 
@@
 Model summaries were generated using the `modelsummary` package [@modelsummary].
+Code style was standardized using `styler` [@styler].
 Code formatting and table presentation were handled by `knitr` [@knitr] and `kableExtra` [@kableExtra], respectively, while regression diagnostics and additional analysis were performed with `car` [@car].
@@
 
-### Model Justification
+## Model Justification
 
@@
 
-### Model Results
+## Model Results
 
@@
 
-## Idealized survey
+## Idealized Survey
 

### 2024-11-28 scripts/00-simulate_data.R

@@ -1,6 +1,6 @@
 #### Preamble ####
-# Purpose: Simulates a dataset to study the impact of several variables, 
-# including education, working hour, region, gender, race and age, on wages. 
+# Purpose: Simulates a dataset to study the impact of several variables,
+# including education, working hour, region, gender, race and age, on wages.
 # Author: Jianing Li
 # Date: 22 November 2024
 # Contact: lijianing.li@mail.utoronto.ca
@@ -30,8 +30,8 @@ region <- sample(c("Northeast", "South", "Midwest", "West"), n_samples, replace 
 # Simulate detailed education levels
 education_level <- sample(
   c("Below_High_School", "High_School", "Some_College", "Bachelor", "Above_Bachelor"),
-  n_samples, 
-  replace = TRUE, 
+  n_samples,
+  replace = TRUE,
   prob = c(0.25, 0.4, 0.2, 0.1, 0.05)
 )
 
@@ -73,13 +73,15 @@ simulated_data <- data.frame(
 
 #  Clean simulated data
 cleaned_data <- simulated_data %>%
-  filter(age >= 18 & age <= 65) %>%  # Keep working age (18-65)
-  filter(UHRSWORK > 0) %>%  # Remove invalid work hours
-  filter(INCWAGE > 0) %>%  # Remove invalid wages
+  filter(age >= 18 & age <= 65) %>% # Keep working age (18-65)
+  filter(UHRSWORK > 0) %>% # Remove invalid work hours
+  filter(INCWAGE > 0) %>% # Remove invalid wages
   mutate(
     region = factor(region),
-    education_level = factor(education_level, levels = c("Below_High_School", "High_School", "Some_College", 
-                                                         "Bachelor", "Above_Bachelor")),
+    education_level = factor(education_level, levels = c(
+      "Below_High_School", "High_School", "Some_College",
+      "Bachelor", "Above_Bachelor"
+    )),
     gender = factor(gender),
     race_group = factor(race_group, levels = c("Asian", "Black", "Other", "White"))
   )
@@ -87,3 +89,4 @@ cleaned_data <- simulated_data %>%
 
 #### Save data ####
 write_csv(cleaned_data, here("data", "00-simulated_data", "simulated_data.csv"))
+

### 2024-11-28 scripts/01-test_simulated_data.R

@@ -4,9 +4,9 @@
 # Date: 22 November 2024
 # Contact: lijianing.li@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: 
-  # - The `tidyverse`, `testthat`, `readr` and `here` packages must be installed and loaded.
-  # - 00-simulate_data.R must have been run
+# Pre-requisites:
+# - The `tidyverse`, `testthat`, `readr` and `here` packages must be installed and loaded.
+# - 00-simulate_data.R must have been run
 # Any other information needed? Make sure you are in the `Determinants-of-Wages` rproj
 
 
@@ -16,44 +16,46 @@ library(testthat)
 library(readr)
 library(here)
 
+#### Test data ####
+
 simulated_data <- read_csv(here("data", "00-simulated_data", "simulated_data.csv"))
 
 
 test_that("Dataset structure and loading", {
   # Test 1: Dataset exists
   expect_true(exists("simulated_data"), "Simulated data was not loaded.")
-  
+
   # Test 2: Dataset is a data frame
   expect_true(is.data.frame(simulated_data), "Simulated data is not a data frame.")
-  
+
   # Test 3: Dataset has non-zero rows and columns
   expect_gt(nrow(simulated_data), 0, "Dataset has no rows.")
   expect_gt(ncol(simulated_data), 0, "Dataset has no columns.")
-  
+
   # Test 4: Dataset contains expected columns
   expected_columns <- c("UHRSWORK", "INCWAGE", "region", "education_level", "age", "gender", "race_group")
-  expect_setequal(names(simulated_data), expected_columns)  # Removed custom error message
+  expect_setequal(names(simulated_data), expected_columns) # Removed custom error message
 })
 
 test_that("Column data types", {
   # Test 5: UHRSWORK is numeric
   expect_type(simulated_data$UHRSWORK, "double")
-  
+
   # Test 6: INCWAGE is numeric
   expect_type(simulated_data$INCWAGE, "double")
-  
+
   # Test 7: region is a character vector
   expect_type(simulated_data$region, "character")
-  
+
   # Test 8: education_level is a character vector
   expect_type(simulated_data$education_level, "character")
-  
+
   # Test 9: age is numeric
   expect_type(simulated_data$age, "double")
-  
+
   # Test 10: gender is a character vector
   expect_type(simulated_data$gender, "character")
-  
+
   # Test 11: race_group is a character vector
   expect_type(simulated_data$race_group, "character")
 })
[…]

### 2024-11-28 scripts/03-clean_data.R

@@ -4,7 +4,7 @@
 # Date: 22 November 2024
 # Contact: lijianing.li@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: 
+# Pre-requisites:
 # - The `tidyverse`, `dplyr`，`arrow` and `here` packages must be installed and loaded.
 # - The steps outlined in 02-download_data.R this document must be followed.
 # Any other information needed? Make sure you are in the `Determinants-of-Wages` rproj
@@ -26,7 +26,7 @@ raw_data <- raw_data %>%
 
 # Clean the data
 cleaned_data <- raw_data %>%
-  filter(!is.na(STATEFIP) & STATEFIP %in% c(1:56)) %>%  # Keep valid state codes
+  filter(!is.na(STATEFIP) & STATEFIP %in% c(1:56)) %>% # Keep valid state codes
   mutate(region = case_when(
     STATEFIP %in% c(9, 23, 25, 33, 34, 36, 42, 44, 50) ~ "Northeast",
     STATEFIP %in% c(17, 18, 19, 20, 26, 27, 29, 31, 38, 39, 46, 55) ~ "Midwest",
@@ -34,19 +34,19 @@ cleaned_data <- raw_data %>%
     STATEFIP %in% c(4, 6, 8, 15, 16, 30, 32, 35, 41, 49, 53, 56) ~ "West"
   )) %>%
   filter(!is.na(region)) %>%
-  filter(!EDUC %in% c(0, 99)) %>%  # Remove invalid education values
+  filter(!EDUC %in% c(0, 99)) %>% # Remove invalid education values
   mutate(education_level = case_when(
-    EDUC %in% c(0, 1, 2, 3, 4, 5) ~ "Below_High_School",  
-    EDUC == 6 ~ "High_School",  
-    EDUC %in% c(7, 8) ~ "Some_College",  
-    EDUC == 10 ~ "Bachelor",  
-    EDUC == 11 ~ "Above_Bachelor"  
+    EDUC %in% c(0, 1, 2, 3, 4, 5) ~ "Below_High_School",
+    EDUC == 6 ~ "High_School",
+    EDUC %in% c(7, 8) ~ "Some_College",
+    EDUC == 10 ~ "Bachelor",
+    EDUC == 11 ~ "Above_Bachelor"
   )) %>%
-  mutate(age = ifelse(AGE == 999, NA, AGE)) %>%  # Replace 999 with NA
[…]

### 2024-11-28 scripts/04-test_analysis_data.R

@@ -4,7 +4,7 @@
 # Date: 22 November 2024
 # Contact: lijianing.li@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: 
+# Pre-requisites:
 # - The `tidyverse`, `testthat`, `readr`，`arrow` and `here` packages must be installed and loaded.
 # - 03-clean_data.R must have been run
 # Any other information needed? Make sure you are in the `Determinants-of-Wages` rproj
@@ -23,14 +23,14 @@ analysis_data <- read_parquet(here("data", "01-analysis_data", "analysis_data.pa
 test_that("Dataset structure and loading", {
   # Test 1: Dataset exists
   expect_true(exists("analysis_data"), "Analysis data was not loaded.")
-  
+
   # Test 2: Dataset is a data frame
   expect_true(is.data.frame(analysis_data), "Analysis data is not a data frame.")
-  
+
   # Test 3: Dataset has non-zero rows and columns
   expect_gt(nrow(analysis_data), 0, "Dataset has no rows.")
   expect_gt(ncol(analysis_data), 0, "Dataset has no columns.")
-  
+
   # Test 4: Dataset contains expected columns
   expected_columns <- c("UHRSWORK", "INCWAGE", "region", "education_level", "age", "gender", "race_group")
   expect_setequal(names(analysis_data), expected_columns)
@@ -40,22 +40,22 @@ test_that("Dataset structure and loading", {
 test_that("Column data types", {
   # Test 5: UHRSWORK is numeric
   expect_type(analysis_data$UHRSWORK, "integer")
-  
+
   # Test 6: INCWAGE is numeric
   expect_type(analysis_data$INCWAGE, "integer")
-  
+
   # Test 7: region is a character vector
   expect_type(analysis_data$region, "character")
-  
+
   # Test 8: education_level is a character vector
   expect_type(analysis_data$education_level, "character")
-  
+
   # Test 9: age is numeric
   expect_type(analysis_data$age, "integer")
-  
+
   # Test 10: gender is a character vector
   expect_type(analysis_data$gender, "character")
-  
+
   # Test 11: race_group is a character vector
   expect_type(analysis_data$race_group, "character")
 })
@@ -64,29 +64,30 @@ test_that("Column data types", {
 test_that("Value ranges and logical consistency", {
   # Test 12: UHRSWORK is strictly greater than 0
[…]

### 2024-11-28 scripts/05-model_data.R

@@ -4,7 +4,7 @@
 # Date: 24 November 2024
 # Contact: lijianing.li@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: 
+# Pre-requisites:
 #- The `tidyverse`， `caret`， `arrow` and `here` packages must be installed and loaded.
 #- 03-clean_data.R must have been run
 # Any other information needed? None
@@ -24,15 +24,15 @@ data <- read_parquet(here("data", "01-analysis_data", "analysis_data.parquet"))
 data <- data %>%
   mutate(log_INCWAGE = log(INCWAGE))
 
-train_index <- createDataPartition(data$log_INCWAGE, p = 0.6, list = FALSE)  
+train_index <- createDataPartition(data$log_INCWAGE, p = 0.6, list = FALSE)
 train_data <- data[train_index, ]
 test_data <- data[-train_index, ]
 
 
 # build a multiple linear regression model that includes a quadratic term for age
 final_model <- lm(
-  formula = log_INCWAGE ~ UHRSWORK + region + education_level + 
-    age + I(age^2) + gender + race_group, 
+  formula = log_INCWAGE ~ UHRSWORK + region + education_level +
+    age + I(age^2) + gender + race_group,
   data = data
 )
 
@@ -54,10 +54,4 @@ saveRDS(
   test_data,
   file = here("models", "test_data.rds")
 )
-
-
-
-
-
-
 

### 2024-12-01 paper/paper.qmd
Final commit

@@
 ---
-title: "Analysing the factors influencing the wages of the US workforce"
+title: "Analysing the Factors Influencing the Wages of the US Workforce in 2023"
 subtitle: "Higher Education and Male Gender as Key Determinants of Wage Increases Controlling for Region and Race"
@@
 The evaluation of the final model's performance was conducted on both training and testing datasets.
-As shown in tbl-test, the RMSE and MAE values are slightly higher for the testing data compared to the training data (0.786 vs.
+As shown in @tbl-test, the RMSE and MAE values are slightly higher for the testing data compared to the training data (0.786 vs.
 0.773 for RMSE, and 0.568 vs.

### 2024-12-03 paper/paper.qmd
Final commit

@@
 workforce, including education, age, gender, race, region, and hours worked.
-The results show that education is the strongest factor, with higher levels of education leading to significantly higher wages, while gender and race reveal notable disparities, with men and Asian workers earning the most.
+The results show that education is the strongest factor, with bachelor’s degree holders earning nearly 50% more than those without a high school diploma, while men earn more than women, and Asian workers have the highest earnings among racial groups.
 The study also finds a nonlinear relationship between age and wages, with region and hours worked playing smaller but significant roles.
@@
 The study finds that education is the strongest determinant of income.
-Higher levels of education lead to significantly higher earnings, with those with a college degree earning much more than other groups.
+Higher levels of education lead to higher earnings, with individuals holding a college degree earning nearly 50% more than those without a high school diploma.
 The relationship between age and earnings is non-linear, with earnings peaking between the ages of 40 and 50 before declining, reflecting the impact of the occupational life cycle.
@@
 These variables cover geographic location, demographic characteristics, education, and employment conditions, providing a broad basis for examining factors influencing wage income.
-The cleaned dataset is presented in @tbl-datasummary, along with detailed descriptions of variable definitions and construction methods, which are thoroughly explained in @sec-data.
+The cleaned dataset is presented in @tbl-datasummary, along with detailed descriptions of variable definitions and construction methods, which are thoroughly explained in @sec-description.
 
@@
 
-Despite its robust sampling methodology and extensive coverage, the ACS data has some inherent limitations.
+Despite its robust sampling methodology and extensive coverage, the ACS data has some limitations.
 Variables such as income and work hours rely heavily on self-reported responses.
@@
 
-This study use Multiple Linear Regression (MLR) to model the relationship between income and various predictors, implemented using R's `lm` function.
-The dataset is divided into training and testing sets using the `createDataPartition` function from the `caret` package, with 60% allocated for model training and parameter estimation and 40% for testing to evaluate predictive performance.
+This study use Multiple Linear Regression (MLR) to model the relationship between income and various predictors, carried out with the `lm` function in R [@citeR].
+The dataset is divided into training and testing sets using the `createDataPartition` function from the `caret` [@caret] package, with 60% allocated for model training and parameter estimation and 40% for testing to evaluate predictive performance.
 
@@
 Gender differences are also evident.
-Male individuals have significantly higher income, as shown by the positive coefficient for gender.
+Male individuals have higher income, as shown by the positive coefficient for gender.
 The narrow confidence interval for gender confirms the reliability of this effect.
 
-Racial and ethnic factors present more complex dynamics.
+Racial factor present more complex dynamics.
 For example, "Black" and "Other" groups have negative coefficients and wider confidence intervals, reflecting greater uncertainty in their effects.
@@
 
-## Dataset Description {#sec-data}
+## Dataset Description {#sec-description}
 

### 2024-12-03 paper/paper.qmd
Final commit

@@
 workforce, including education, age, gender, race, region, and hours worked.
-The results show that education is the strongest factor, with bachelor’s degree holders earning nearly 50% more than those without a high school diploma, while men earn more than women, and Asian workers have the highest earnings among racial groups.
+The results show that education is the strongest factor, with individuals holding a bachelor’s degree earning over twice as much as those without a high school diploma, while men earn more than women, and Asian workers have the highest earnings among racial groups.
 The study also finds a nonlinear relationship between age and wages, with region and hours worked playing smaller but significant roles.
@@
 
-Chart (g) shows education levels: 16.5% of the population has less than a high school education, 30.4% has completed high school, 23.5% has some college education, 26.2% holds a bachelor’s degree, and 9.5% has education beyond a bachelor’s degree.
-Chart (h) highlights the connection between education and income.
-Those with less than a high school education earn $36,149 on average, while individuals with a bachelor’s degree or higher earn $87,465 and $124,144, respectively.
-Income increases steadily with higher levels of education.
+Chart (g) illustrates the distribution of education levels: 16.5% of the population has less than a high school education, 30.4% has completed high school, 23.5% has some college education, 26.2% holds a bachelor’s degree, and 9.5% has education beyond a bachelor’s degree.
+Chart (h) highlights the strong relationship between education and income.
+Those with less than a high school education earn an average of $36,149, while individuals with a bachelor’s degree earn $87,465, which is approximately 2.4 times higher.
+Income continues to increase for those with education beyond a bachelor’s degree, averaging $124,144, further underscoring the significant impact of education on earnings.
 

### 2024-12-03 paper/paper.qmd
Final commit

@@
 This study use Multiple Linear Regression (MLR) to model the relationship between income and various predictors, carried out with the `lm` function in R [@citeR].
-The dataset is divided into training and testing sets using the `createDataPartition` function from the `caret` [@caret] package, with 60% allocated for model training and parameter estimation and 40% for testing to evaluate predictive performance.
+The dataset is divided into training and testing sets using the `createDataPartition` function from the `caret` package [@caret], with 60% allocated for model training and parameter estimation and 40% for testing to evaluate predictive performance.
 

### 2024-12-06 paper/paper.qmd
Final commit

@@
 Additionally, existing research often lacks recent data that are both nationally representative and adaptable for diverse analyses.
-This study addresses these gaps using the 2023 sample from the IPUMS database, which allows users to customize variables and provides high-quality, comprehensive data.
-Leveraging this resource, the study identifies the main determinants of income inequality through a multiple linear regression model.
+This study addresses these gaps using the 2023 sample from the IPUMS database, which allows users to customize variables and provides high-quality, detailed data.
+Using this resource, the study identifies the main determinants of income inequality through a multiple linear regression model.
 
@@
 
-This paper provides a comprehensive analysis of income inequality by examining its key drivers and impacts.
+This paper provides a thorough analysis of income inequality by examining its main factors and effects.
 @sec-data describes the dataset in detail, highlights its main features, and uses visualizations to present the analysis results.
@@
 It also identifies the study’s limitations and suggests directions for future research.
-@sec-appendix provides additional details on the model, and methods of survey and sampling, offering more comprehensive support for the study.
+@sec-appendix provides additional details on the model, and methods of survey and sampling, offering more detailed support for the study.
 
@@
 The sample includes individuals living in private households and those in group quarters, such as student dormitories, correctional facilities, and care homes.
-Weights are applied to the data to reduce biases caused by the sampling design and nonresponse.
-This improves the generalizability of the results.
+Weights are applied to reduce biases from the sampling design and nonresponse, enhancing result generalizability.
 The smallest geographic unit in the dataset is the Public Use Microdata Area (PUMA).
@@
 
-The ACS collects data using multiple methods.
-These include mailed questionnaires, online submissions, and in-person or telephone interviews.
+The ACS collects data through mailed questionnaires, online submissions, and in-person or telephone interviews.
 The surveys cover a wide range of topics, such as demographics, education, employment, and income.
@@
 
-Despite its robust sampling methodology and extensive coverage, the ACS data has some limitations.
-Variables such as income and work hours rely heavily on self-reported responses.
-This introduces potential inaccuracies due to recall bias, rounding errors, or deliberate misreporting.
-For individuals with irregular or multiple income sources, questions about wages or hours worked may be interpreted differently, further increasing measurement errors.
+Despite its robust sampling methodology and broad coverage, the ACS data has limitations.
+Variables such as income and work hours rely heavily on self-reported responses, introducing potential inaccuracies due to recall bias, rounding errors, or deliberate misreporting.
+For individuals with irregular or multiple income sources, questions about wages or hours worked may be interpreted variably, increasing measurement errors.
 
@@
 However, these methods cannot fully eliminate the risk of bias or error.
-Even with these limitations, the ACS dataset remains a reliable and widely used resource for studying wage income and related factors.
+Despite these limitations, the ACS dataset remains a reliable resource for analyzing wage income and related factors.
 Its combination of sampling weights, large sample size, and rigorous design framework supports robust analysis.
@@
 @fig-demographics-distribution illustrates the distribution of demographic characteristics and their relationship with average income across gender, race, region, and education level.
-Pie charts display the proportional distribution of each group, while bar charts show corresponding average income levels, providing a straightforward depiction of these relationships.
-
-Chart (a) shows the gender distribution: females account for 48.9% of the population, and males 51.1%, indicating a nearly equal split.
-Chart (b) shows that males earn an average income of $82,840, compared to $57,146 for females.
+Group proportions are shown alongside their corresponding average income levels, clearly illustrating these relationships.
+
+Chart (a) illustrates gender distribution: females make up 48.9% and males 51.1% of the population, indicating a near-equal split.
+Chart (b) shows males with an average income of $82,840, compared to $57,146 for females.
 
 Chart (c) presents the racial composition, with White individuals forming the largest group at 67%, followed by Black individuals at 18.2%, Asians at 6.8%, and other racial groups at 8.1%.
-Chart (d) shows that Asians have the highest average income at $96,574, followed by White individuals at $73,778.
+Chart (d) shows Asians with the highest average income at $96,574, followed by White individuals at $73,778.
 Black individuals and other racial groups earn $51,221 and $56,604, respectively.
 
-Chart (e) shows the regional distribution.
+Chart (e) illustrates regional distribution.
 The West represents the largest share at 29%, followed by the Northeast (26.6%), the South (24.1%), and the Midwest (20.3%).
@@
 
-Chart (g) illustrates the distribution of education levels: 16.5% of the population has less than a high school education, 30.4% has completed high school, 23.5% has some college education, 26.2% holds a bachelor’s degree, and 9.5% has education beyond a bachelor’s degree.
+Chart (g) illustrates education levels: 16.5% have less than a high school education, 30.4% completed high school, 23.5% attended some college, 26.2% hold a bachelor’s degree, and 9.5% pursued education beyond a bachelor’s degree.
 Chart (h) highlights the strong relationship between education and income.
-Those with less than a high school education earn an average of $36,149, while individuals with a bachelor’s degree earn $87,465, which is approximately 2.4 times higher.
-Income continues to increase for those with education beyond a bachelor’s degree, averaging $124,144, further underscoring the significant impact of education on earnings.
+Those with less than a high school education earn an average of $36,149, while individuals with a bachelor’s degree earn $87,465, approximately 2.4 times higher.
+Income rises further for those with education beyond a bachelor’s degree, averaging $124,144, highlighting the substantial impact of education on earnings.
 
@@
 ```
-
 @fig-combined-visualizations examines the relationship between income, work hours, and age, incorporating both individual-level data and aggregated trends.
 
-Chart (a) visualizes the relationship between work hours and income, where the red quadratic regression line indicates a positive correlation.
-Longer work hours generally correspond to higher income, with the regression line reflecting a stable linear trend rather than noticeable deceleration.
-Income variability increases with longer work hours, particularly beyond 50 hours per week, where fluctuations become more pronounced.
-Chart (b) represents the trend of average income by work hours.
-Income shows a steady increase with work hours, especially between 20 and 50 hours per week, but becomes less consistent beyond 50 hours.
-
-Chart (c) illustrates the relationship between age and income, with the red quadratic regression line highlighting a pattern where income rises with age before gradually declining.
-Chart (d) presents average income by age as a line chart.
-It shows that average income steadily increases with age, peaking in the late 40s and early 50s, followed by a gradual decline as individuals approach retirement age.
+Chart (a) shows the relationship between work hours and income, with the red quadratic regression line indicating a positive correlation.
+Longer work hours generally correspond to higher income, with the regression line reflecting a stable linear trend without noticeable deceleration.
+Income variability grows with longer work hours, especially beyond 50 hours per week, where fluctuations are more pronounced.
+Chart (b) depicts average income trends across work hours.
+Income steadily increases with work hours, especially between 20 and 50 hours per week, but becomes less consistent beyond 50 hours.
+
+Chart (c) shows the relationship between age and income, with the red quadratic regression line revealing a rise in income with age before a gradual decline.
+Chart (d) shows average income by age as a line graph.
+It reveals that average income steadily increases with age, peaking in the late 40s and early 50s, followed by a gradual decline as individuals approach retirement age.
 
@@
 
-@fig-Education-Race illustrates the proportion of different education levels across racial groups. The intensity of blue indicates the proportion, with darker shades representing higher values.
-
-Among Asians, the proportion of highly educated individuals is relatively high, with 38% holding a bachelor's degree and 30% having education beyond a bachelor's degree, while the proportion of those with "Below High School" education is 0%.
-In the Black group, "High School" accounts for the largest proportion at 32%, followed by "Some College" at 27%, while those with "Above Bachelor" degrees make up only 16%.
-In the White group, "High School" accounts for 29%, with 27% holding a bachelor's degree and 24% having "Some College" education, while 17% have "Above Bachelor" degrees.
-Among the "Other" group, 40% have a "High School" education, 25% have "Some College," 17% hold a bachelor's degree, and only 9% have "Above Bachelor" degrees.
-
-Educational attainment shows significant variation across racial groups, reflecting the complex relationship between race and educational background.
-Asians have the highest levels of education, with a strong representation in bachelor's and advanced degrees.
-In contrast, Black and White groups display a broader distribution, with high school education being more common.
-The "Other" category, encompassing smaller racial groups and mixed-race individuals, stands out with a concentration at the high school level and lower representation in higher education, suggesting potential socio-economic and cultural disparities.
+@fig-Education-Race illustrates the proportion of different education levels across racial groups. Darker shades of blue indicate higher proportions.
+
+Among Asians, 38% hold a bachelor's degree and 30% have education beyond it, while none fall under "Below High School." For Black individuals, "High School" is most common at 32%, followed by "Some College" at 27%, and "Above Bachelor" at 16%.
+In the White group, 29% have a "High School" education, 27% hold a bachelor's degree, 24% have "Some College" education, and 17% have "Above Bachelor" degrees.
+The "Other" category shows 40% with a "High School" education, 25% with "Some College," 17% with a bachelor's degree, and only 9% with "Above Bachelor" degrees.
+
+Educational attainment varies significantly across racial groups, reflecting the nuanced relationship between race and education.
+Asians exhibit the highest levels of education, with strong representation in bachelor's and advanced degrees.
+In contrast, Black and White groups show broader distributions, with high school education being more common.
+The "Other" category, which includes smaller racial groups and mixed-race individuals, is concentrated at the high school level and underrepresented in higher education, indicating possible socio-economic and cultural disparities.
 
@@
 
-The goal of our modelling strategy is to quantify the contributions of key factors—such as weekly working hours, age (and its quadratic term), education level, region, gender, and race—to variations in income (log-transformed).
-By estimating the relative impact of these predictors, the model provides insights into potential social and demographic disparities in income.
+The goal of this modelling strategy is to quantify the contributions of factors such as weekly working hours, age (including its quadratic term), education level, region, gender, and race to variations in log-transformed income.
+By estimating the relative impact of these predictors, the model examines potential social and demographic disparities in income.
 
@@
 
-Multiple regression models rely on several assumptions: linearity, meaning a linear relationship exists between independent and dependent variables; homoskedasticity, requiring constant error variance across predictors; independence of errors, meaning residuals are uncorrelated; normality, where residuals follow a normal distribution; and independence of independent variables, ensuring no significant multicollinearity.
+Multiple regression models rely on several assumptions: linearity, indicating a linear relationship between predictors and the dependent variable; homoskedasticity, requiring constant error variance across predictors; independence of errors, where residuals are uncorrelated; normality, where residuals follow a normal distribution; and low multicollinearity among predictors.
 These assumptions are discussed in @sec-model-details and evaluated using the diagnostic plots shown in @fig-check.
@@
 The selection of variables and model structure was based on theory and data characteristics.
-Given the right-skewed nature of the income distribution in @fig-income-distribution, the dependent variable is log-transformed (`log(INCWAGE)`) to improve model fit.
-Furthermore, to capture the nonlinear relationship between age and income evident in @fig-combined-visualizations, a quadratic term for age (`age^2`) is included in the model.
+To address the right-skewed income distribution shown in @fig-income-distribution, the dependent variable is log-transformed (log(INCWAGE)) to improve model fit.
+Additionally, a quadratic term for age (age^2) is included to capture the nonlinear relationship between age and income observed in @fig-combined-visualizations.
 Categorical variables, such as region, education level, race group, and gender, are represented using dummy variables.
-Each category is compared to a reference category, allowing the regression model to capture and interpret group-level differences.
+Each category is compared to a reference group, enabling the model to interpret group-level differences.
 
@@
 This interaction term aimed to evaluate income differences across racial groups at the same education level.
-However, most interaction terms were not statistically significant and provided minimal improvement to model fit metrics, such as adjusted $R^2$ and AIC.
+However, most interaction terms were not statistically significant and contributed minimally to metrics like adjusted $R^2$ and AIC.
 To simplify the model and enhance interpretability, this interaction term was excluded from the final specification.
 
-The evaluation of the final model's performance was conducted on both training and testing datasets.
+The final model's performance was evaluated on both training and testing datasets.
 As shown in @tbl-test, the RMSE and MAE values are slightly higher for the testing data compared to the training data (0.786 vs.
@@
 0.543 for MAE), reflecting a modest decline in predictive accuracy when applied to unseen data.
-Similarly, the $R^2$ values are 0.488 for the training set and 0.468 for the testing set, indicating that the model captures a moderate proportion of the variability in the log-transformed income.
-Additionally, all variables included in the final model were statistically significant, with $p$-values below 0.05, further supporting the robustness of the selected predictors.
+Similarly, the $R^2$ values of 0.488 for the training set and 0.468 for the testing set indicate that the model captures a moderate share of the variability in log-transformed income.
+Additionally, all variables included in the final model were statistically significant, with $p$-values below 0.05, supporting the robustness of the selected predictors.
 Overall, the final model achieves a practical balance between complexity, interpretability, and predictive performance.
@@
 @fig-coefficients shows the effects of weekly working hours, age, region, education level, gender, and racial group on income.
-The coefficient plot also displays the 95% confidence intervals for each coefficient, with detailed numerical values presented in @tbl-modelresults in @sec-model-details.
+The plot displays 95% confidence intervals for each coefficient, with detailed numerical values provided in @tbl-modelresults in @sec-model-details.
 
@@
 Individuals with lower education, such as "Below High School," have negative coefficients, while those with higher education, such as a "Bachelor’s" degree, show positive coefficients.
-This highlights the significant economic benefits of higher education.
+This underscores the economic advantages of higher education.
 However, the confidence intervals for education vary.
@@
 
-Age also has a strong impact on income.
+Age significantly influences income.
 The positive coefficient for age shows that income increases as people get older.
@@
 
-Racial factor present more complex dynamics.
+Racial factors present more complex dynamics.
 For example, "Black" and "Other" groups have negative coefficients and wider confidence intervals, reflecting greater uncertainty in their effects.
@@
 Its narrow confidence interval indicates a stable and reliable estimate.
-Overall, these results provide a clear understanding of the socioeconomic factors influencing income and offer valuable insights for future research and policy discussions.
+Overall, these results clarify the socioeconomic factors influencing income and offer meaningful implications for future research and policy discussions.
 
@@
 
-This paper uses a multiple linear regression model to analyze key factors that influence income, including education, age, gender, race, region, and hours worked.
-The study is based on data from the 2023 American Community Survey (ACS), cleaned and processed for visualization and analysis.
-The model quantifies the impact of these variables and explores their significance, showing how personal characteristics and external factors shape the income distribution.
+This paper uses a multiple linear regression model to analyze factors influencing income, including education, age, gender, race, region, and hours worked.
+This study uses data from the 2023 American Community Survey (ACS), which was cleaned and processed for analysis.
+The model quantifies the impact of these variables and explores their significance, showing how personal characteristics and external factors shape income distribution.
 Education, age, gender, and race are treated as endogenous variables, reflecting individual attributes, while region and work hours are considered exogenous, capturing the influence of external conditions on income.
-Through the analysis of these variables, the study provides insights into the roots of income inequality and offers evidence to inform policymaking.
-
-## Education and Age: Core Drivers of Income
-
-Education is a key factor influencing income, with a clear positive correlation between education level and earnings.
+By analyzing these variables, the study provides evidence to inform policymaking and examines the roots of income inequality.
+
+## Education and Age: Primary Determinants of Income
+
+Education strongly influences income, with a clear positive correlation between education level and earnings.
 Graduate degree holders earn significantly more, while a bachelor's degree also boosts income.
 However, those with only 1-2 years of college earn similar incomes to high school graduates, underscoring the importance of completing college.
-Higher education plays a crucial role in enhancing economic potential and securing high-paying jobs.
+Higher education enhances economic potential and secures high-paying jobs.
 Most income differences stem from variations in bachelor's, graduate, and doctoral education levels, while pre-college education has a smaller impact.
 
-Age also has a nonlinear impact on income.
+Age significantly affects income.
 The model indicates that income peaks between ages 40 and 50 before gradually declining.
-The quadratic term captures this trend, reflecting how different stages of a career affect earnings.
-Early in their careers, individuals experience rapid income growth as they gain experience and promotions.
+The quadratic term captures this trend, reflecting how career stages impact earnings.
+Early in their careers, individuals see rapid income growth with experience and promotions.
 In later stages, productivity declines, and income typically decreases as retirement approaches.
@@
 Education and age shape an individual's earning potential.
-These findings are valuable for career planning and offer guidance for policymakers.
+These findings inform career planning and provide guidance for policymakers.
 Individuals can enhance their competitiveness in the job market by pursuing a bachelor's degree or higher.
@@
 
-Gender inequality in income distribution is evident.
+Income distribution shows clear gender inequality.
 Studies show that men earn significantly more than women, even when controlling for education, age, and other factors.
@@
 
-Racial disparities, on the other hand, highlight another form of structural inequality.
+Racial disparities underscore another form of structural inequality.
 Research indicates that Asians earn the highest incomes, followed by Whites, while Black individuals and those from "other" racial groups earn the least.
@@
 
-Gender and race are exogenous factors, beyond individual control, and the wage disparities they cause reflect systemic inequality.
+Gender and race are exogenous factors beyond individual control.
 Addressing the gender pay gap requires targeted policies, such as promoting pay transparency and increasing women's participation in high-paying industries.
@@
 
-Work hours are positively linked to income, but the returns diminish.
+Income increases with work hours, but returns diminish.
 Beyond 50 hours per week, income becomes more unstable.
@@
 
-The data used in this study has certain limitations.
+This study's data has certain limitations.
 It relies on survey responses, which can include self-reporting bias.
@@
 This exclusion might have introduced bias into the results.
-To simplify the model, smaller racial groups, such as Indigenous, and multiracial individuals, were categorized as “Other.” Although this made the analysis manageable, it overlooked finer details in racial differences.
+To simplify the model, smaller racial groups, such as Indigenous, and multiracial individuals, were categorized as "Other." While this approach simplified the analysis, it overlooked finer racial differences.
 Future studies could refine these categories and include more detailed variables when sample sizes allow.
